---
title: "BeWell - Sensitivity Analysis"
format: 
  html:
    code-fold: TRUE
editor_options: 
  chunk_output_type: console
---
Last updated: 4/23/2026

# Load packages

```{r load_packages}
if (!require(pacman)) install.packages("pacman")
library(pacman)
## Load/install/update the packages
p_load(tidyverse, 
       janitor, 
       pivottabler, 
       psych, 
       summarytools, 
       lme4,
       lmerTest,
       MuMIn,
       scales,
       ggplot2,
       ggrepel,
       ggeffects,
       sjPlot,
       splines,
       multcomp,
       effsize,
       naniar,
       mice,
       miceadds,
       mitools,
       mitml,
       MBESS,
       broom.mixed,
       stats,
       reshape2,
       ggcorrplot,
       gt,
       gtsummary,
       gtExtras,
       showtext,
       patchwork,
       flextable,
       brms, 
       VGAM,
       ggpubr, 
       corrplot,
       performance,
       glmmTMB,
       DHARMa)
options(scipen = 0)
```

# Functions

## Function: Mapping prefix to variable names
```{r fx_prefix_titles}
map_prefix <- function(var_name) {
  target_prefix <- list(
    phq9_sum = "sr",
    phq8_sum = "sr",
    gad7_total = "sr",
    flourishing_flourishing = "sr",
    promisSleep_tScore = "sr",
    medTherapy_counseling = "sr",
    medTherapy_medAntiDep = "sr",
    medTherapy_medOther = "sr",
    medTherapy_bin = "sr",
    mst_LDI = "beh",
    mst_correctedRecognition = "beh",
    linguistic_negative_sentiment = "vlg",
    linguistic_i = "vlg",
    crp_dbs_mean_AB = "dbs",
    crp_plasma_mean_AB = "dbs",
    cyto_il6_dbs_mean_AB = "dbs",
    Clostridium_IV_clr = "mib",
    Desulfovibrio_clr = "mib",
    Faecalibacterium_clr = "mib",
    Enterobacteriaceae_clr = "mib",
    shannon_genus_estimate = "mib"
  )

  cleaned <- gsub("^(sr|dbs|mib|vlg|beh|sea)_(base|scr|pw1|pw2|pw3|pw4|3mo)_", "", var_name)
  
  match_name <- names(target_prefix)[
    startsWith(cleaned, names(target_prefix))  # take first match
  ]

  if (length(match_name) > 0) {
    return(target_prefix[[match_name[1]]])
  } else {
    message("No Prefix Found for: ", var_name)
    return(NA)
  }
}
```

## Function: Mapping worse direction to variable names

```{r fx_map_worse_direction}
map_worse <- function(var_name) {
  worse_direction <- c(phq8_sum = "H",
                     gad7_total = "H",
                     promisSleep_tScore = "H",
                     flourishing_flourishing = "L",
                     linguistic_negative_sentiment = "H",
                     linguistic_i = "L",
                     mst_LDI = "L",
                     mst_correctedRecognition = "L",
                     linguistic_positive_sentiment = "L",
                     cheekRaise_full_mean = "L",
                     Duchenne_full_mean_log = 'L',
                     crp_dbs_mean_AB = "H",
                     crp_plasma_mean_AB = "H",
                     cyto_tnfAlpha_dbs_mean_AB = "H",
                     cyto_il6_dbs_mean_AB = "H",
                     cyto_il10_dbs_mean_AB = "H",
                     Clostridium_IV_clr = "H",
                     Desulfovibrio_clr = "L",
                     Faecalibacterium_clr = "H",
                     Enterobacteriaceae_clr = "L",
                     shannon_genus_estimate = "H",
                     simpson_genus_estimate_log = "H")
  
  if (var_name %in% names(worse_direction)) {
    return(worse_direction[[var_name]])
  } else {
    cleaned <- gsub("^(sr|dbs|mib|vlg|beh|sea)_(base|scr|pw1|pw2|pw3|pw4|3mo)_", "", var_name)
  
    if (cleaned %in% names(worse_direction)) {
      return(worse_direction[[cleaned]])
    } else {
      print('Variable not found')
    }
  }
}
```

## Function: Mapping titles

```{r title_list}
target_titles <- list(
  phq8_sum = "PHQ-8",
  gad7_total = "GAD-7",
  promisSleep_tScore = "PROMIS Sleep Disturbance",
  flourishing_flourishing = "Flourishing Index",
  crp_plasma_mean_AB = "C-Reactive Protein",
  cyto_il6_dbs_mean_AB = "Interleukin-6",
  shannon_genus_estimate = "Alpha Diversity",
  Enterobacteriaceae_clr = "Enterobacteriaceae",
  Desulfovibrio_clr = "Desulfovibrio",
  Faecalibacterium_clr = "Faecalibacterium",
  Clostridium_IV_clr = "Clostridium IV",
  mst_LDI = "Lure Discrimination Index",
  mst_correctedRecognition = "Recognition Memory",
  linguistic_negative_sentiment = "Negative Affect Words",
  linguistic_i = "First-Person Pronouns",
  medTherapy_counseling = "Received professional counseling or psychotherapy to support mental health in the past month",
  medTherapy_medAntiDep = "Currently taking antidepressant medications",
  medTherapy_medOther = "Took other prescribed medications to support mental health in the past month",
  medTherapy_bin = "Recent mental health treatment (therapy or medication use)"
)

map_title <- function(var_name, title_map) {
  if (var_name %in% names(title_map)) {
    return(title_map[[var_name]])
  } else {
    cleaned <- gsub("^(sr|dbs|mib|vlg|beh|sea)_(base|scr|pw1|pw2|pw3|pw4|3mo)_", "", var_name)
  
    if (cleaned %in% names(title_map)) {
      return(title_map[[cleaned]])
    } else {
      return(var_name)
    }
  }
}

map_title_with_baseline <- function(x, title_map) {
  # Ensure input is a single string
  if (!is.character(x) || length(x) != 1) {
    stop("x must be a single string")
  }
  
  if (str_detect(x, "_base_")) {
    return("Baseline")
  } else {
    return(map_title(x, title_map))
  }
}
```

## Function: Converting dfs from wide format to long format

```{r wide_to_long_fx}
convert_long_format <- function(df_tmp, target_var_strings, constant_vars){
  
  cat("Looking for columns ending with:", paste(target_var_strings, collapse=", "), "\n")
  
  ids <- df_tmp$id
  
  df_tmp_long <- crossing(id = ids, time = 1:6) %>% 
    mutate(week = case_when(
      time == 1 ~ 0,
      time == 2 ~ 1,
      time == 3 ~ 2,
      time == 4 ~ 3,
      time == 5 ~ 4,
      time == 6 ~ 17
    ),
    time1 = recode_values(time,
                       1 ~ -4,
                       2 ~ -3,
                       3 ~ -2,
                       4 ~ -1,
                       c(5, 6) ~ 0),
    time2 = recode_values(time,
                       c(1, 2, 3, 4, 5) ~ 0,
                       6 ~ 1),
    uniqueid = paste0(id, "_", time))
  
  df_tmp_long <- df_tmp_long %>%
    left_join(df_tmp %>% 
                dplyr::select('id', all_of(constant_vars)), by = "id")
  
  for (i in seq_along(target_var_strings)) {
    target_var <- target_var_strings[i]
    print(target_var)

    df_i <- df_tmp %>%
      pivot_longer(cols = all_of(ends_with(target_var)), 
                   names_to = "var",
                   values_to = target_var,
                   values_drop_na = TRUE) %>% 
      dplyr::select(id, sr_scr_rand_group, var, !!sym(target_var)) %>%
      mutate(time = case_when(
        (substring(var, 4, 6) == 'bas' | substring(var, 5, 7) == 'bas') ~ 1,
        (substring(var, 4, 6) == 'pw1' | substring(var, 5, 7) == 'pw1') ~ 2,
        (substring(var, 4, 6) == 'pw2' | substring(var, 5, 7) == 'pw2') ~ 3,
        (substring(var, 4, 6) == 'pw3' | substring(var, 5, 7) == 'pw3') ~ 4,
        (substring(var, 4, 6) == 'pw4' | substring(var, 5, 7) == 'pw4') ~ 5,
        (substring(var, 4, 6) == '3mo' | substring(var, 5, 7) == '3mo') ~ 6
      ),
      !!sym(target_var) := as.numeric(!!sym(target_var)),
      uniqueid = paste0(id, "_", time)) %>% 
      dplyr::select(uniqueid, !!sym(target_var))

    df_tmp_long <- left_join(df_tmp_long, df_i, by = "uniqueid")
  }
  
  df_tmp_long <- df_tmp_long %>% 
    dplyr::select(-uniqueid)
  
  return(df_tmp_long)
}
```

## Function: General linear hypothesis test for long format data

```{r glht_fx}
glht_models <- function(df_tmp, out_vars, mi_results = "N", ref_level = 'hmp_full', return_results = FALSE) {
  
  if (length(out_vars) > 1){
    
    models <- vector("list", length(out_vars) * 2)
    
    for (i in seq_along(out_vars)) {
      outcome_var <- out_vars[i]
      i1 <- i * 2 - 1
      i2 <- i * 2
      formula <- as.formula(paste(outcome_var, "~ time1*sr_scr_rand_group + time2*sr_scr_rand_group + (time1|id)"))
      
      if (ref_level == 'hmp_full') {
        analysis_info1 <- glht(lmer(formula, data = df_tmp),
                               linfct = c("time1:sr_scr_rand_grouphmp_active*4 + sr_scr_rand_grouphmp_active:time2 = 0"))
        analysis_info2 <- glht(lmer(formula, data = df_tmp),
                               linfct = c("time1:sr_scr_rand_groupuc*4 + sr_scr_rand_groupuc:time2 = 0"))
      } else if (ref_level == 'hmp_active') {
        analysis_info1 <- glht(lmer(formula, data = df_tmp),
                               linfct = c("time1:sr_scr_rand_grouphmp_full*4 + sr_scr_rand_grouphmp_full:time2 = 0"))
        analysis_info2 <- glht(lmer(formula, data = df_tmp),
                               linfct = c("time1:sr_scr_rand_groupuc*4 + sr_scr_rand_groupuc:time2 = 0"))
      }
      
      models[[i1]] <- analysis_info1
      models[[i2]] <- analysis_info2
      
      cat("\nModel:", i1, " - ", outcome_var, " full vs active:\n")
      print(summary(analysis_info1))
      
      if (ref_level == 'hmp_full') {
        cat("\nModel:", i2, " - ", outcome_var, " full vs uc:\n")
      } else if (ref_level == 'hmp_active'){
        cat("\nModel:", i2, " - ", outcome_var, " active vs uc:\n")
      }
      print(summary(analysis_info2))
    }
    
    if (return_results == TRUE){
      return(models)
    }
    
  } else {
      outcome_var <- out_vars
      formula <- as.formula(paste(outcome_var, "~ time1*sr_scr_rand_group + time2*sr_scr_rand_group + (time1|id)"))
      
      if (ref_level == 'hmp_full') {
        analysis_info1 <- glht(lmer(formula, data = df_tmp),
                               linfct = c("time1:sr_scr_rand_grouphmp_active*4 + sr_scr_rand_grouphmp_active:time2 = 0"))
        analysis_info2 <- glht(lmer(formula, data = df_tmp),
                               linfct = c("time1:sr_scr_rand_groupuc*4 + sr_scr_rand_groupuc:time2 = 0"))
      } else if (ref_level == 'hmp_active') {
        analysis_info1 <- glht(lmer(formula, data = df_tmp),
                               linfct = c("time1:sr_scr_rand_grouphmp_full*4 + sr_scr_rand_grouphmp_full:time2 = 0"))
        analysis_info2 <- glht(lmer(formula, data = df_tmp),
                               linfct = c("time1:sr_scr_rand_groupuc*4 + sr_scr_rand_groupuc:time2 = 0"))
      }
      
      models <- list(analysis_info1, analysis_info2)
      
      cat("\nModel: 1 - ", outcome_var, " full vs active:\n")
      print(summary(analysis_info1))
      
      if (ref_level == 'hmp_full') {
        cat("\nModel: 2 - ", outcome_var, " full vs uc:\n")
      } else if (ref_level == 'hmp_active'){
        cat("\nModel: 2 - ", outcome_var, " active vs uc:\n")
      }
      print(summary(analysis_info2))
      
    if (return_results == TRUE){
      return(models)
    }
  }
  
}
```

## Function: GLHT for imputed data (long format)

```{r fx_glht_imputed}

glht_models_mi <- function(outcome_var, fit_list, return_results = FALSE) {
  L_active <- "time1:sr_scr_rand_grouphmp_active*4 + sr_scr_rand_grouphmp_active:time2 = 0"
  L_uc <- "time1:sr_scr_rand_groupuc*4 + sr_scr_rand_groupuc:time2 = 0"
  
  active_estimates <- vector("numeric", length(fit_list$analyses))
  active_std_errors <- vector("numeric", length(fit_list$analyses))
  uc_estimates <- vector("numeric", length(fit_list$analyses))
  uc_std_errors <- vector("numeric", length(fit_list$analyses))
  
  for (j in seq_along(fit_list$analyses)) {
    model <- fit_list$analyses[[j]]
    
    active_test <- glht(model, linfct = L_active)
    active_summary <- summary(active_test)
    active_estimates[j] <- active_summary$test$coefficients
    active_std_errors[j] <- active_summary$test$sigma
    
    uc_test <- glht(model, linfct = L_uc)
    uc_summary <- summary(uc_test)
    uc_estimates[j] <- uc_summary$test$coefficients
    uc_std_errors[j] <- uc_summary$test$sigma
  }
  
  m <- length(fit_list$analyses)
  
  active_est_pooled <- mean(active_estimates)
  active_var_within <- mean(active_std_errors^2)
  active_var_between <- sum((active_estimates - active_est_pooled)^2) / (m - 1)
  active_var_total <- active_var_within + (1 + 1/m) * active_var_between
  
  # Calculate degrees of freedom using Barnard-Rubin adjustment
  active_df <- (m - 1) * (1 + active_var_within / ((1 + 1/m) * active_var_between))^2
  active_t <- active_est_pooled / sqrt(active_var_total)
  active_p <- 2 * pt(abs(active_t), df = active_df, lower.tail = FALSE)
  active_ci_lower <- active_est_pooled - qt(0.975, df = active_df) * sqrt(active_var_total)
  active_ci_upper <- active_est_pooled + qt(0.975, df = active_df) * sqrt(active_var_total)
  
  uc_est_pooled <- mean(uc_estimates)
  uc_var_within <- mean(uc_std_errors^2)
  uc_var_between <- sum((uc_estimates - uc_est_pooled)^2) / (m - 1)
  uc_var_total <- uc_var_within + (1 + 1/m) * uc_var_between
  
  uc_df <- (m - 1) * (1 + uc_var_within / ((1 + 1/m) * uc_var_between))^2
  uc_t <- uc_est_pooled / sqrt(uc_var_total)
  uc_p <- 2 * pt(abs(uc_t), df = uc_df, lower.tail = FALSE)
  uc_ci_lower <- uc_est_pooled - qt(0.975, df = uc_df) * sqrt(uc_var_total)
  uc_ci_upper <- uc_est_pooled + qt(0.975, df = uc_df) * sqrt(uc_var_total)
  
  cat("\nModel:", outcome_var, " full vs active (pooled):\n")
  cat("Estimate:", round(active_est_pooled, 4), 
      "Std. Error:", round(sqrt(active_var_total), 4),
      "t value:", round(active_t, 4), 
      "df:", round(active_df, 2),
      "Pr(>|t|):", round(active_p, 4), "\n")
  
  cat("\nModel:", outcome_var, " full vs uc (pooled):\n")
  cat("Estimate:", round(uc_est_pooled, 4), 
      "Std. Error:", round(sqrt(uc_var_total), 4),
      "t value:", round(uc_t, 4),
      "df:", round(uc_df, 2),
      "Pr(>|t|):", round(uc_p, 4), "\n")
  
    if (return_results) {
    return(list(
      active = list(
        estimate = active_est_pooled,
        ci_lower = active_ci_lower,
        ci_upper = active_ci_upper,
        p_value = active_p
      ),
      uc = list(
        estimate = uc_est_pooled,
        ci_lower = uc_ci_lower,
        ci_upper = uc_ci_upper,
        p_value = uc_p
      )
    ))
  }
}

```

## Function: General linear hypothesis test for moderation (long format data)

```{r glht_fx_moderation}
glht_models_moderation <- function(df_tmp, out_vars, moderator_var, mi_results = "N", return_results = FALSE) {
  
  if (length(out_vars) > 1){
    
    models <- vector("list", length(out_vars) * 2)
    
    for (i in seq_along(out_vars)) {
      outcome_var <- out_vars[i]
      i1 <- i * 2 - 1
      i2 <- i * 2
      formula <- as.formula(paste(outcome_var, "~ time1*sr_scr_rand_group*", moderator_var, " + time2*sr_scr_rand_group*", moderator_var, " + (time1|id)"))
      
      
      
      analysis_info1 <- glht(lmer(formula, data = df_tmp),
                             linfct = c(paste0("time1:sr_scr_rand_grouphmp_active:", moderator_var, "*4 + sr_scr_rand_grouphmp_active:", moderator_var, ":time2 = 0")))
      analysis_info2 <- glht(lmer(formula, data = df_tmp),
                             linfct = c(paste0("time1:sr_scr_rand_groupuc:", moderator_var, "*4 + sr_scr_rand_groupuc:", moderator_var, ":time2 = 0")))
      
      models[[i1]] <- analysis_info1
      models[[i2]] <- analysis_info2
      
      cat("\nModel:", i1, " - ", outcome_var, " full vs active:\n")
      print(summary(analysis_info1))
      
      cat("\nModel:", i2, " - ", outcome_var, " full vs uc:\n")
      print(summary(analysis_info2))
    }
  } else {
      outcome_var <- out_vars
      formula <- as.formula(paste(outcome_var, "~ time1*sr_scr_rand_group*", moderator_var, " + time2*sr_scr_rand_group*", moderator_var, " + (time1|id)"))
      
      analysis_info1 <- glht(lmer(formula, data = df_tmp),
                             linfct = c(paste0("time1:sr_scr_rand_grouphmp_active:", moderator_var, "*4 + sr_scr_rand_grouphmp_active:", moderator_var, ":time2 = 0")))
      analysis_info2 <- glht(lmer(formula, data = df_tmp),
                             linfct = c(paste0("time1:sr_scr_rand_groupuc:", moderator_var, "*4 + sr_scr_rand_groupuc:", moderator_var, ":time2 = 0")))
      
      models <- list(analysis_info1, analysis_info2)
      
      cat("\nModel: 1 - ", outcome_var, " full vs active:\n")
      print(summary(analysis_info1))
      
      cat("\nModel: 2 - ", outcome_var, " full vs uc:\n")
      print(summary(analysis_info2))
      
    if (return_results == TRUE){
      return(models)
    }
  }
  
}
```

## Function: Effect size computation for long format data

```{r fx_effect_size_imputed}
compute_effect_size_long <- function(estimation, data, var, preMI = c('Y', 'N'), og_df, ref_level = 'hmp_full'){
  preMI <- match.arg(preMI)
  
  if (ref_level == 'hmp_full'){
    group1_t1_term <- "time1:sr_scr_rand_grouphmp_active"
    group1_t2_term <- "sr_scr_rand_grouphmp_active:time2"
  } else if (ref_level == 'hmp_active'){
    group1_t1_term <- "time1:sr_scr_rand_grouphmp_full"
    group1_t2_term <- "sr_scr_rand_grouphmp_full:time2"
  }
  
  if (preMI == 'Y') {
    sum_coef <- summary(estimation)$coefficients
    
    group1_t1_est <- sum_coef[group1_t1_term, "Estimate"]
    group1_t2_est <- sum_coef[group1_t2_term, "Estimate"]
    uc_t1_est <- sum_coef["time1:sr_scr_rand_groupuc", "Estimate"]
    uc_t2_est <- sum_coef["sr_scr_rand_groupuc:time2", "Estimate"]
  } else if (preMI == 'N') {
    group1_t1_est <- summary(estimation) %>% 
      filter(term == group1_t1_term) %>% 
      pull(estimate)
    group1_t2_est <-  summary(estimation) %>% 
      filter(term == group1_t2_term) %>% 
      pull(estimate)
    uc_t1_est <- summary(estimation) %>% 
      filter(term =="time1:sr_scr_rand_groupuc") %>% 
      pull(estimate)
    uc_t2_est <- summary(estimation) %>% 
      filter(term =="sr_scr_rand_groupuc:time2") %>% 
      pull(estimate)
  }
  
  var_prefix <- map_prefix(var)
  
  base_var <- ifelse(str_detect(var, '_min0'), 
                 paste0(var_prefix, '_base_', str_remove(var, '_min0')), 
                 paste0(var_prefix, '_base_', var))
  
  #stat from original df~~~~~~~~~~~~~
  

  sd_base_full <- sd((og_df %>%
                        filter(sr_scr_rand_group %in% c('hmp_full')) %>%
                        dplyr::select(!!sym(base_var)) %>% 
                        unlist()),
                     na.rm = TRUE)
  sd_base_active <- sd((og_df %>%
                        filter(sr_scr_rand_group %in% c('hmp_active')) %>%
                        dplyr::select(!!sym(base_var)) %>% 
                        unlist()),
                     na.rm = TRUE)
  sd_base_uc <- sd((og_df %>% 
                          filter(sr_scr_rand_group %in% c('uc')) %>%
                          dplyr::select(!!sym(base_var)) %>% 
                          unlist()),
                       na.rm = TRUE)

    
  # Convert mids to long format if needed for sample size calculation
  if (inherits(data, "mids")) {
    data_long <- mice::complete(data, action = "long", include = TRUE)
  } else {
    data_long <- data
  }
  
  # sample sizes from actual data~~~~~~~~~~~~~
  t1_n <- data_long %>% 
    filter(!is.na(!!sym(var)),
            time <= 5) %>% 
    group_by(sr_scr_rand_group) %>% 
    summarize(unique_id = n_distinct(id), .groups = 'drop')
  
  full_t1_n <- t1_n[t1_n$sr_scr_rand_group == 'hmp_full', 'unique_id'] %>% unlist()
  active_t1_n <- t1_n[t1_n$sr_scr_rand_group == 'hmp_active', 'unique_id'] %>% unlist()
  uc_t1_n <- t1_n[t1_n$sr_scr_rand_group == 'uc', 'unique_id'] %>% unlist()
  
  t1t2_n <- data_long %>% 
    filter(!is.na(!!sym(var))) %>% 
    group_by(sr_scr_rand_group) %>% 
    summarize(unique_id = n_distinct(id), .groups = 'drop')
  
  full_t1t2_n <- t1t2_n[t1t2_n$sr_scr_rand_group == 'hmp_full', 'unique_id'] %>% unlist()
  active_t1t2_n <- t1t2_n[t1t2_n$sr_scr_rand_group == 'hmp_active', 'unique_id'] %>% unlist()
  uc_t1t2_n <- t1t2_n[t1t2_n$sr_scr_rand_group == 'uc', 'unique_id'] %>% unlist()
  
  # SDs
  sd_base_activeFull <- sqrt(((full_t1_n - 1)*sd_base_full^2 + (active_t1_n - 1)*sd_base_active^2) / (full_t1_n + active_t1_n - 2))
  sd_base_ucFull <- sqrt(((full_t1_n - 1)*sd_base_full^2 + (uc_t1_n - 1)*sd_base_uc^2) / (full_t1_n + uc_t1_n - 2))
  sd_base_ucActive <- sqrt(((active_t1_n - 1)*sd_base_active^2 + (uc_t1_n - 1)*sd_base_uc^2) / (active_t1_n + uc_t1_n - 2))
  
  #~~~~~~~~Between-group effect size~~~~~~~~~~~~~~~~~~~
  
  full_active_smd_t1 <- (group1_t1_est * 4) / sd_base_activeFull
  full_active_ci_t1 <- ci.smd(smd = full_active_smd_t1, n.1 = full_t1_n, n.2 = active_t1_n)
  full_active_smd_t1t2 <- (group1_t1_est * 4 + group1_t2_est * 1) / sd_base_activeFull
  full_active_ci_t1t2 <- ci.smd(smd = full_active_smd_t1t2, n.1 = full_t1t2_n, n.2 = active_t1t2_n)
  
  if (ref_level == 'hmp_full'){
    group1_uc_smd_t1 <- (uc_t1_est * 4) / sd_base_ucFull
    group1_uc_ci_t1 <- ci.smd(smd = group1_uc_smd_t1, n.1 = full_t1_n, n.2 = uc_t1_n)
    group1_uc_smd_t1t2 <- (uc_t1_est * 4 + uc_t2_est * 1) / sd_base_ucFull
    group1_uc_ci_t1t2 <- ci.smd(smd = group1_uc_smd_t1t2, n.1 = full_t1t2_n, n.2 = uc_t1t2_n)
  } else if (ref_level == 'hmp_active'){
    group1_uc_smd_t1 <- (uc_t1_est * 4) / sd_base_ucActive
    group1_uc_ci_t1 <- ci.smd(smd = group1_uc_smd_t1, n.1 = active_t1_n, n.2 = uc_t1_n)
    group1_uc_smd_t1t2 <- (uc_t1_est * 4 + uc_t2_est * 1) / sd_base_ucActive
    group1_uc_ci_t1t2 <- ci.smd(smd = group1_uc_smd_t1t2, n.1 = active_t1t2_n, n.2 = uc_t1t2_n)
  }
  
  cat('\n******Between-group effect size ******\n')
  cat('\nIntervention vs Active Ctrl******\n')
  cat('time 1 x group Estimate:', group1_t1_est, '\n')
  cat('time 2 x group Estimate:', group1_t2_est, '\n')
  cat('Intervention Time 1 sample size:', full_t1_n, '\n')
  cat('Active ctrl Time 1 sample size:', active_t1_n, '\n')
  cat('Intervention overall sample size:', full_t1t2_n, '\n')
  cat('Active ctrl overall sample size:', active_t1t2_n, '\n')
  cat('post-test effect size:\n', 
      '(', group1_t1_est, '* 4) /', sd_base_activeFull, '=', full_active_smd_t1, '\n')
  cat('95% CI:', round(full_active_ci_t1$Lower, 4), ',', round(full_active_ci_t1$Upper, 4), '\n')
  cat('overall effect size:\n', 
      '(', group1_t1_est, '* 4 +', group1_t2_est, '* 1) /', sd_base_activeFull, '=', full_active_smd_t1t2, '\n')
  cat('95% CI:', round(full_active_ci_t1t2$Lower, 4), ',', round(full_active_ci_t1t2$Upper, 4), '\n')
  
  if (ref_level == 'hmp_full'){
    cat('\nIntervention vs Usual care******\n')
    cat('time 1 x group Estimate:', uc_t1_est, '\n')
    cat('time 2 x group Estimate:', uc_t2_est, '\n')
    cat('Intervention Time 1 sample size:', full_t1_n, '\n')
    cat('Usual care Time 1 sample size:', uc_t1_n, '\n')
    cat('Intervention overall sample size:', full_t1t2_n, '\n')
    cat('Usual care overall sample size:', uc_t1t2_n, '\n')
    cat('post-test effect size:\n', 
        '(', uc_t1_est, '* 4) /', sd_base_ucFull, '=', group1_uc_smd_t1, '\n')
    cat('95% CI:', round(group1_uc_ci_t1$Lower, 4), ',', round(group1_uc_ci_t1$Upper, 4), '\n')
    cat('overall effect size:\n', 
        '(', uc_t1_est, '* 4 +', uc_t2_est, '* 1) /', sd_base_ucFull, '=', group1_uc_smd_t1t2, '\n')
    cat('95% CI:', round(group1_uc_ci_t1t2$Lower, 4), ',', round(group1_uc_ci_t1t2$Upper, 4), '\n\n')
    group_comparisons <- c("full_vs_active", "full_vs_uc")
  } else if (ref_level == 'hmp_active'){
    cat('\nActive Ctrl vs Usual care******\n')
    cat('time 1 x group Estimate:', uc_t1_est, '\n')
    cat('time 2 x group Estimate:', uc_t2_est, '\n')
    cat('Active ctrl Time 1 sample size:', active_t1_n, '\n')
    cat('Usual care Time 1 sample size:', uc_t1_n, '\n')
    cat('Active ctrl overall sample size:', active_t1t2_n, '\n')
    cat('Usual care overall sample size:', uc_t1t2_n, '\n')
    cat('post-test effect size:\n', 
        '(', uc_t1_est, '* 4) /', sd_base_ucActive, '=', group1_uc_smd_t1, '\n')
    cat('95% CI:', round(group1_uc_ci_t1$Lower, 4), ',', round(group1_uc_ci_t1$Upper, 4), '\n')
    cat('overall effect size:\n', 
        '(', uc_t1_est, '* 4 +', uc_t2_est, '* 1) /', sd_base_ucActive, '=', group1_uc_smd_t1t2, '\n')
    cat('95% CI:', round(group1_uc_ci_t1t2$Lower, 4), ',', round(group1_uc_ci_t1t2$Upper, 4), '\n\n')
    group_comparisons <- c("active_vs_full", "active_vs_uc")
  }
  
  
  
  # Return all the effect size as a data frame
  effect_df <- data.frame(
    group = c(
      group_comparisons[1], group_comparisons[1],
      group_comparisons[2], group_comparisons[2]
    ),
    effect_type = c(
      "T1", "Overall",
      "T1", "Overall"
    ),
    d = c(
      full_active_smd_t1,
      full_active_smd_t1t2,
      group1_uc_smd_t1,
      group1_uc_smd_t1t2
    ),
    ci_lower = c(
      full_active_ci_t1$Lower,
      full_active_ci_t1t2$Lower,
      group1_uc_ci_t1$Lower,
      group1_uc_ci_t1t2$Lower
    ),
    ci_upper = c(
      full_active_ci_t1$Upper,
      full_active_ci_t1t2$Upper,
      group1_uc_ci_t1$Upper,
      group1_uc_ci_t1t2$Upper
    )
  )
  
  return(effect_df)
}
```

## Function: Effect size computation with moderator for long format data

```{r fx_effect_size_imputed_moderation}
compute_effect_size_long_moderation <- function(estimation, data, var, moderator_var, preMI = c('Y', 'N'), og_df){
  preMI <- match.arg(preMI)
  
  time1_active_mod <- paste0("time1:sr_scr_rand_grouphmp_active:", moderator_var)
  time2_active_mod <- paste0("sr_scr_rand_grouphmp_active:", moderator_var, ":time2")
  time1_uc_mod <- paste0("time1:sr_scr_rand_groupuc:", moderator_var)
  time2_uc_mod <- paste0("sr_scr_rand_groupuc:", moderator_var, ":time2")
  
  
  if (preMI == 'Y') {
    sum_coef <- summary(estimation)$coefficients
    
    active_t1_est <- sum_coef[time1_active_mod, "Estimate"]
    active_t2_est <- sum_coef[time2_active_mod, "Estimate"]
    uc_t1_est <- sum_coef[time1_uc_mod, "Estimate"]
    uc_t2_est <- sum_coef[time2_uc_mod, "Estimate"]
  } else if (preMI == 'N') {
    active_t1_est <- summary(estimation) %>% 
      filter(term == time1_active_mod) %>% 
      pull(estimate)
    active_t2_est <-  summary(estimation) %>% 
      filter(term ==time2_active_mod) %>% 
      pull(estimate)
    uc_t1_est <- summary(estimation) %>% 
      filter(term == time1_uc_mod) %>% 
      pull(estimate)
    uc_t2_est <- summary(estimation) %>% 
      filter(term == time2_uc_mod) %>% 
      pull(estimate)
  }
  
  var_prefix <- map_prefix(var)
  
  base_var <- ifelse(str_detect(var, '_min0'), 
                 paste0(var_prefix, '_base_', str_remove(var, '_min0')), 
                 paste0(var_prefix, '_base_', var))
  
  #stat from original df~~~~~~~~~~~~~
  
  full_active_sd_base <- sd((og_df %>% 
                          filter(sr_scr_rand_group %in% c('hmp_full', 'hmp_active')) %>%
                          dplyr::select(!!sym(base_var)) %>% unlist()),
                       na.rm = TRUE)
  
  full_uc_sd_base <- sd((og_df %>% 
                          filter(sr_scr_rand_group %in% c('hmp_full', 'uc')) %>%
                          dplyr::select(!!sym(base_var)) %>% unlist()),
                       na.rm = TRUE)
    
  # Convert mids to long format if needed for sample size calculation
  if (inherits(data, "mids")) {
    data_long <- mice::complete(data, action = "long", include = TRUE)
  } else {
    data_long <- data
  }
  
  # sample sizes from actual data~~~~~~~~~~~~~
  t1_n <- data_long %>% 
    filter(!is.na(!!sym(var)),
            time <= 5) %>% 
    group_by(sr_scr_rand_group) %>% 
    summarize(unique_id = n_distinct(id), .groups = 'drop')
  
  full_t1_n <- t1_n[t1_n$sr_scr_rand_group == 'hmp_full', 'unique_id'] %>% unlist()
  active_t1_n <- t1_n[t1_n$sr_scr_rand_group == 'hmp_active', 'unique_id'] %>% unlist()
  uc_t1_n <- t1_n[t1_n$sr_scr_rand_group == 'uc', 'unique_id'] %>% unlist()
  
  t1t2_n <- data_long %>% 
    filter(!is.na(!!sym(var))) %>% 
    group_by(sr_scr_rand_group) %>% 
    summarize(unique_id = n_distinct(id), .groups = 'drop')
  
  full_t1t2_n <- t1t2_n[t1t2_n$sr_scr_rand_group == 'hmp_full', 'unique_id'] %>% unlist()
  active_t1t2_n <- t1t2_n[t1t2_n$sr_scr_rand_group == 'hmp_active', 'unique_id'] %>% unlist()
  uc_t1t2_n <- t1t2_n[t1t2_n$sr_scr_rand_group == 'uc', 'unique_id'] %>% unlist()
  
  #~~~~~~~~Between-group effect size~~~~~~~~~~~~~~~~~~~
    
  # Adding the moderator mean in effect size calculation 
  moderator_mean <- mean(og_df[[moderator_var]], na.rm = TRUE)
  
  full_active_smd_t1 <- (moderator_mean*(active_t1_est * 4)) / full_active_sd_base
  full_active_ci_t1 <- ci.smd(smd = full_active_smd_t1, n.1 = full_t1_n, n.2 = active_t1_n)
  full_active_smd_t1t2 <- (moderator_mean*(active_t1_est * 4 + active_t2_est * 1)) / full_active_sd_base
  full_active_ci_t1t2 <- ci.smd(smd = full_active_smd_t1t2, n.1 = full_t1t2_n, n.2 = active_t1t2_n)
  
  full_uc_smd_t1 <- (moderator_mean*(uc_t1_est * 4)) / full_uc_sd_base
  full_uc_ci_t1 <- ci.smd(smd = full_uc_smd_t1, n.1 = full_t1_n, n.2 = uc_t1_n)
  full_uc_smd_t1t2 <- (moderator_mean*(uc_t1_est * 4 + uc_t2_est * 1)) / full_uc_sd_base
  full_uc_ci_t1t2 <- ci.smd(smd = full_uc_smd_t1t2, n.1 = full_t1t2_n, n.2 = uc_t1t2_n)

  
  cat('\n******Mean of', moderator_var, '******\n')
  cat('\n', moderator_mean, '\n')
  
  cat('\n******Between-group effect size ******\n')
  cat('\nIntervention vs Active Ctrl******\n')
  cat('time 1 x group x ', moderator_var, ' Estimate:', active_t1_est, '\n')
  cat('time 2 x group x ', moderator_var, ' Estimate:', active_t2_est, '\n')
  cat('Intervention Time 1 sample size:', full_t1_n, '\n')
  cat('Active ctrl Time 2 sample size:', active_t1_n, '\n')
  cat('Intervention overall sample size:', full_t1t2_n, '\n')
  cat('Active ctrl overall sample size:', active_t1t2_n, '\n')
  cat('post-test effect size:\n', 
      '(', moderator_mean, '*(', active_t1_est, '* 4) /', full_active_sd_base, ') =', full_active_smd_t1, '\n')
  cat('95% CI:', round(full_active_ci_t1$Lower, 4), ',', round(full_active_ci_t1$Upper, 4), '\n')
  cat('overall effect size:\n', 
      '(', moderator_mean, '*(', active_t1_est, '* 4 +', active_t2_est, '* 1) /', full_active_sd_base, ') =', full_active_smd_t1t2, '\n')
  cat('95% CI:', round(full_active_ci_t1t2$Lower, 4), ',', round(full_active_ci_t1t2$Upper, 4), '\n')
  
  cat('\nIntervention vs Usual care******\n')
  cat('time 1 x group x ', moderator_var, ' Estimate:', uc_t1_est, '\n')
  cat('time 2 x group x ', moderator_var, ' Estimate:', uc_t2_est, '\n')
  cat('Intervention Time 1 sample size:', full_t1_n, '\n')
  cat('Usual care Time 2 sample size:', uc_t1_n, '\n')
  cat('Intervention overall sample size:', full_t1t2_n, '\n')
  cat('Usual care overall sample size:', uc_t1t2_n, '\n')
  cat('post-test effect size:\n', 
      '(', moderator_mean, '*(', uc_t1_est, '* 4) /', full_uc_sd_base, ') =', full_uc_smd_t1, '\n')
  cat('95% CI:', round(full_uc_ci_t1$Lower, 4), ',', round(full_uc_ci_t1$Upper, 4), '\n')
  cat('overall effect size:\n', 
      '(', moderator_mean, '*(', uc_t1_est, '* 4 +', uc_t2_est, '* 1) /', full_uc_sd_base, ') =', full_uc_smd_t1t2, '\n')
  cat('95% CI:', round(full_uc_ci_t1t2$Lower, 4), ',', round(full_uc_ci_t1t2$Upper, 4), '\n\n')
  
  # Return all the effect size as a data frame
  effect_df <- data.frame(
    group = c(
      "full_vs_active", "full_vs_active",
      "full_vs_uc", "full_vs_uc"
    ),
    effect_type = c(
      "T1", "Overall",
      "T1", "Overall"
    ),
    d = c(
      full_active_smd_t1,
      full_active_smd_t1t2,
      full_uc_smd_t1,
      full_uc_smd_t1t2
    ),
    ci_lower = c(
      full_active_ci_t1$Lower,
      full_active_ci_t1t2$Lower,
      full_uc_ci_t1$Lower,
      full_uc_ci_t1t2$Lower
    ),
    ci_upper = c(
      full_active_ci_t1$Upper,
      full_active_ci_t1t2$Upper,
      full_uc_ci_t1$Upper,
      full_uc_ci_t1t2$Upper
    )
  )
  
  return(effect_df)
}
```

## Function: Scale imputed values

If higher level = worse, then the scaling should be done with 1.1 and 1.2, while if lower level = worse, then it should be done with 0.8 and 0.9

```{r fx_scale_imputed}
scale_imputed_values_long <- function(df_tmp, outcome) {
  worse <- map_worse(outcome)
  
  missing_var <- paste0("missing_", outcome)
  
  if (worse == "H"){
    var110 <- paste0(outcome, "_110")
    var120 <- paste0(outcome, "_120")
    
    df_tmp <- df_tmp %>% 
      mutate(!!sym(var110) := if_else(!!sym(missing_var) == 1, 
                                      !!sym(outcome) * 1.1,
                                      !!sym(outcome)),
             !!sym(var120) := if_else(!!sym(missing_var) == 1, 
                                      !!sym(outcome) * 1.2,
                                      !!sym(outcome))
             )
  } else if (worse == "L"){
    var80 <- paste0(outcome, "_80")
    var90 <- paste0(outcome, "_90")
    
    df_tmp <- df_tmp %>% 
      mutate(!!sym(var80) := if_else(!!sym(missing_var) == 1, 
                                     !!sym(outcome) * 0.8,
                                     !!sym(outcome)),
             !!sym(var90) := if_else(!!sym(missing_var) == 1, 
                                     !!sym(outcome) * 0.9,
                                     !!sym(outcome))
             )
  }
  
  print(tail(df_tmp %>% filter(!!sym(missing_var) == 1)))
  
  return(df_tmp)
}
```


## Function: Piecewise model with imputed data

OG model

```{r fx_piecewise_original}
run_piecewise_og_long <- function(outcome, covar_formula = NULL, og_df_long = df_long, og_df_wide = df, covariates = NULL, return_table = FALSE, exc_outlier = FALSE, ref_level = 'hmp_full'){
  
  og_df_long <- og_df_long %>% 
    mutate(sr_scr_rand_group = relevel(factor(sr_scr_rand_group), ref = ref_level))
  
  if (!is.null(covariates)){
    covar_formula <- paste0(covariates, collapse=" + ")
    covar_formula <- paste0(covar_formula, "+ ")
  } else {
    covar_formula <- NULL
  }
  
  if (exc_outlier == TRUE){ #exc outlier (mean +/- >3SDs) if necessary, this will overrides og_df_wide and og_df_long!!!
    og_df_wide <- og_df_wide %>% 
      mutate(across(
        matches(outcome),
        ~ {
          x <- .
          threshold_upper <- mean(x, na.rm = TRUE) + 3 * sd(x, na.rm = TRUE)
          threshold_lower <- mean(x, na.rm = TRUE) - 3 * sd(x, na.rm = TRUE)
          replace(x, x >= threshold_upper | x <= threshold_lower, NA)
        }
      ))
    
    og_df_long <- convert_long_format(df_tmp = og_df_wide, 
                                      target_var_strings = outcome,
                                   constant_vars = 'sr_scr_rand_group')
  }
  
  og_formula <- as.formula(paste(outcome, "~ ", covar_formula, "time1*sr_scr_rand_group + time2*sr_scr_rand_group + (time1|id)"))
  
  cat('\n\n')
  print(og_formula)
  og_est <- lmer(og_formula, data = og_df_long)
  summary_og_est <- summary(og_est)
  fixed_effects <- summary_og_est$coefficients
  fixed_effects_rounded <- round(fixed_effects, 4)

  confint_og_est <- confint(og_est)
  confint_fixed <- confint_og_est[rownames(confint_og_est) %in% rownames(fixed_effects), ]
  confint_fixed_rounded <- round(confint_fixed, 4)

  results_table <- cbind(
    Estimate = fixed_effects_rounded[, "Estimate"],
    Std_Error = fixed_effects_rounded[, "Std. Error"],
    t_value = fixed_effects_rounded[, "t value"],
    df = fixed_effects_rounded[, "df"],
    p_value = fixed_effects_rounded[, "Pr(>|t|)"],
    `CI_2.5%` = confint_fixed_rounded[, 1],
    `CI_97.5%` = confint_fixed_rounded[, 2]
    )

  print(results_table)

  print(r.squaredGLMM(og_est))
  
  cat('\n****** Effect Size summary ******\n')
  effect_df <- compute_effect_size_long(estimation = og_est,
                        data = og_df_long,
                        var = outcome,
                        preMI = 'Y',
                        og_df = og_df_wide,
                        ref_level = ref_level)
  
  
  glht_models_all <- glht_models(og_df_long, as.character(outcome), return_results = TRUE, ref_level = ref_level)
  
  if (return_table == TRUE){
  glht_full_active <- glht_models_all[[1]]
  glht_full_active_summary <- summary(glht_full_active)
  glht_group1_uc <- glht_models_all[[2]]
  glht_group1_uc_summary <- summary(glht_group1_uc)
  
  if (ref_level == 'hmp_full'){
    group_comparison <- c("full_vs_active", "full_vs_uc")
    group_comparison_cap <- c("Full_vs_Active", "Full_vs_UC")
    group1_time1_term <- "time1:sr_scr_rand_grouphmp_active"
  } else if (ref_level == 'hmp_active'){
    group_comparison <- c("active_vs_full", "active_vs_uc")
    group_comparison_cap <- c("Active_vs_Full", "Active_vs_UC")
    group1_time1_term <- "time1:sr_scr_rand_grouphmp_full"
  }
    
  glht_table <- data.frame(
    comparison = group_comparison,
    estimate = c(coef(glht_full_active_summary)[1], coef(glht_group1_uc_summary)[1]),
    t_value = c(glht_full_active_summary$test$tstat[1], glht_group1_uc_summary$test$tstat[1]),
    p_value = c(glht_full_active_summary$test$pvalues[1], glht_group1_uc_summary$test$pvalues[1]),
    ci_lower = c(confint(glht_full_active)$confint[1, "lwr"], confint(glht_group1_uc)$confint[1, "lwr"]),
    ci_upper = c(confint(glht_full_active)$confint[1, "upr"], confint(glht_group1_uc)$confint[1, "upr"])
  )
  
   summary_table <- data.frame(
    outcome = outcome,
    comparison = rep(group_comparison_cap, each = 2),
    timepoint = rep(c("Week 4 vs. Baseline", "Week 17 vs. Baseline"), 2),
    
    # Coefficients
    coef = c(
      fixed_effects[group1_time1_term, "Estimate"],
      coef(glht_full_active_summary)[1],
      fixed_effects["time1:sr_scr_rand_groupuc", "Estimate"],
      coef(glht_group1_uc_summary)[1]
    ),
    
    # Coefficient CIs
    coef_ci_lower = c(
      confint_fixed[group1_time1_term, 1],
      confint(glht_full_active)$confint[1, "lwr"],
      confint_fixed["time1:sr_scr_rand_groupuc", 1],
      confint(glht_group1_uc)$confint[1, "lwr"]
    ),
    coef_ci_upper = c(
      confint_fixed[group1_time1_term, 2],
      confint(glht_full_active)$confint[1, "upr"],
      confint_fixed["time1:sr_scr_rand_groupuc", 2],
      confint(glht_group1_uc)$confint[1, "upr"]
    ),
    
    # P-values
    p_value = c(
      fixed_effects[group1_time1_term, "Pr(>|t|)"],
      glht_full_active_summary$test$pvalues[1],
      fixed_effects["time1:sr_scr_rand_groupuc", "Pr(>|t|)"],
      glht_group1_uc_summary$test$pvalues[1]
    ),
    
    # Effect sizes
    d = effect_df$d,
    d_ci_lower = effect_df$ci_lower,
    d_ci_upper = effect_df$ci_upper,
    
    stringsAsFactors = FALSE
  )
   
     # Round numeric columns
  summary_table <- summary_table %>%
    mutate(across(where(is.numeric), ~round(., 4)))
  
    return(list(model = og_est, results_summary = summary_table, ci_fixed = confint_fixed))
  }
}
```

OG model with moderator

```{r fx_piecewise_original_moderation}
run_piecewise_og_long_moderation <- function(outcome, covar_formula = NULL, og_df_long = df_long, og_df_wide = df, moderator_var, covariates = NULL, return_table = FALSE){
  if (!is.null(covariates)){
    covar_formula <- paste0(covariates, collapse=" + ")
    covar_formula <- paste0(covar_formula, "+ ")
  } else {
    covar_formula <- NULL
  }

  og_formula <- as.formula(paste(outcome, "~ ", covar_formula, "time1*sr_scr_rand_group*", moderator_var, " + time2*sr_scr_rand_group*", moderator_var, " + (time1|id)"))
  
  
  cat('\n\n')
  print(og_formula)
  og_est <- lmer(og_formula, data = og_df_long)
  summary_og_est <- summary(og_est)
  fixed_effects <- summary_og_est$coefficients
  fixed_effects_rounded <- round(fixed_effects, 4)

  confint_og_est <- confint(og_est)
  confint_fixed <- confint_og_est[rownames(confint_og_est) %in% rownames(fixed_effects), ]
  confint_fixed_rounded <- round(confint_fixed, 4)

  results_table <- cbind(
    Estimate = fixed_effects_rounded[, "Estimate"],
    Std_Error = fixed_effects_rounded[, "Std. Error"],
    t_value = fixed_effects_rounded[, "t value"],
    df = fixed_effects_rounded[, "df"],
    p_value = fixed_effects_rounded[, "Pr(>|t|)"],
    `CI_2.5%` = confint_fixed_rounded[, 1],
    `CI_97.5%` = confint_fixed_rounded[, 2]
    )

  print(results_table)

  print(r.squaredGLMM(og_est))
  
  cat('\n****** Effect Size summary ******\n')
    effect_df <- compute_effect_size_long_moderation(estimation = og_est,
                                                     data = og_df_long,
                                                     var = outcome,
                                                     moderator_var = moderator_var,
                                                     preMI = 'Y',
                                                     og_df = og_df_wide)
    
    
    
  if (return_table == TRUE){
      glht_models_all <- glht_models_moderation(og_df_long, 
                                                as.character(outcome), 
                                                moderator_var = moderator_var,
                                                return_results = TRUE)
      glht_full_active <- glht_models_all[[1]]
      glht_full_active_summary <- summary(glht_full_active)
      glht_full_uc <- glht_models_all[[2]]
      glht_full_uc_summary <- summary(glht_full_uc)
      
      glht_table <- data.frame(
        comparison = c("full_vs_active", "full_vs_uc"),
        estimate = c(coef(glht_full_active_summary)[1], coef(glht_full_uc_summary)[1]),
        t_value = c(glht_full_active_summary$test$tstat[1], glht_full_uc_summary$test$tstat[1]),
        p_value = c(glht_full_active_summary$test$pvalues[1], glht_full_uc_summary$test$pvalues[1]),
        ci_lower = c(confint(glht_full_active)$confint[1, "lwr"], confint(glht_full_uc)$confint[1, "lwr"]),
        ci_upper = c(confint(glht_full_active)$confint[1, "upr"], confint(glht_full_uc)$confint[1, "upr"])
      )
  
      time1_active_mod <- paste0("time1:sr_scr_rand_grouphmp_active:", moderator_var)
      time1_uc_mod <- paste0("time1:sr_scr_rand_groupuc:", moderator_var)
      time2_active_mod <- paste0("sr_scr_rand_grouphmp_active:", moderator_var, ":time2")
      time2_uc_mod <- paste0("sr_scr_rand_groupuc:", moderator_var, ":time2")
      
       summary_table <- data.frame(
        outcome = outcome,
        comparison = rep(c("Full_vs_Active", "Full_vs_UC"), each = 2),
        timepoint = rep(c("Week 4 vs. Baseline", "Week 17 vs. Baseline"), 2),
        
        # Coefficients
        coef = c(
          fixed_effects[time1_active_mod, "Estimate"],
          coef(glht_full_active_summary)[1],
          fixed_effects[time1_uc_mod, "Estimate"],
          coef(glht_full_uc_summary)[1]
        ),
        
        # Coefficient CIs
        coef_ci_lower = c(
          confint_fixed[time1_active_mod, 1],
          confint(glht_full_active)$confint[1, "lwr"],
          confint_fixed[time1_uc_mod, 1],
          confint(glht_full_uc)$confint[1, "lwr"]
        ),
        coef_ci_upper = c(
          confint_fixed[time1_active_mod, 2],
          confint(glht_full_active)$confint[1, "upr"],
          confint_fixed[time1_uc_mod, 2],
          confint(glht_full_uc)$confint[1, "upr"]
        ),
        
        # P-values
        p_value = c(
          fixed_effects[time1_active_mod, "Pr(>|t|)"],
          glht_full_active_summary$test$pvalues[1],
          fixed_effects[time1_uc_mod, "Pr(>|t|)"],
          glht_full_uc_summary$test$pvalues[1]
        ),
        
        # Effect sizes
        d = effect_df$d,
        d_ci_lower = effect_df$ci_lower,
        d_ci_upper = effect_df$ci_upper,
        
        stringsAsFactors = FALSE
      )
       
     # Round numeric columns
  summary_table <- summary_table %>%
    mutate(across(where(is.numeric), ~round(., 4)))
  
    return(list(model = og_est, results_summary = summary_table, ci_fixed = confint_fixed))
  }
}
```

## Function for imputed piecewise models
There are four models:
- Original model
- Model with imputed data
- Model with imputed data, in which the data is scaled to be 10% worse
- Model with imputed data, in which the data is scaled to be 20% worse

```{r run_sensitivity_imputed_long}
run_sensitivity_imputed_long <- function(df_tmp, outcome, 
                                         covariates = NULL, og_df_wide = df, 
                                         og_df_long = df_long, return_table = FALSE,
                                         return_models = FALSE){
  
  worse <- map_worse(outcome)
  
  if (!is.null(covariates)){
    covar_formula <- paste0(covariates, collapse=" + ")
    covar_formula <- paste0(covar_formula, "+ ")
  } else {
    covar_formula <- NULL
  }

  # Original model (unchanged)
  cat("\nOutcome:", outcome,"(Original, NOT imputed) -----------------------\n")
  orig_results <- run_piecewise_og_long(outcome,
                          covar_formula,
                          og_df_wide = og_df_wide,
                          og_df_long = og_df_long,
                          return_table = TRUE)
  
  og_summary <- orig_results$results_summary
  og_summary <- og_summary %>% 
    mutate(model_type = "Original") %>% 
    relocate(outcome, model_type)

  mids_tmp <- mice::as.mids(df_tmp)
  
  process_model <- function(outcome_name, model_label) {
    cat("\nOutcome:", outcome_name, "(", model_label, ") -----------------------\n")
    
    model_check <- with(mids_tmp,
      lmer(formula(paste(outcome_name, "~ ", covar_formula, "time1*sr_scr_rand_group + time2*sr_scr_rand_group + (time1|id)")),
           control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))
    )
    
    est <- pool(model_check)
    summary_est <- summary(est, conf.int = TRUE)
    summary_est[, 2:ncol(summary_est)] <- round(summary_est[, 2:ncol(summary_est)], 4)
    print(summary_est)
    
    effect_df <- compute_effect_size_long(estimation = est, data = mids_tmp, 
                                          var = outcome, preMI = 'N', og_df = og_df_wide)
    
    glht_results <- glht_models_mi(outcome_name, model_check, return_results = return_table)
    
    if (return_table) {
      time1_active_idx <- which(summary_est$term == "time1:sr_scr_rand_grouphmp_active")
      time1_uc_idx <- which(summary_est$term == "time1:sr_scr_rand_groupuc")
      
      summary_table <- data.frame(
        outcome = outcome,
        model_type = model_label,
        comparison = rep(c("Full_vs_Active", "Full_vs_UC"), each = 2),
        timepoint = rep(c("Week 4 vs. Baseline", "Week 17 vs. Baseline"), 2),
        coef = c(summary_est$estimate[time1_active_idx], glht_results$active$estimate,
                summary_est$estimate[time1_uc_idx], glht_results$uc$estimate),
        coef_ci_lower = c(summary_est$`2.5 %`[time1_active_idx], glht_results$active$ci_lower,
                         summary_est$`2.5 %`[time1_uc_idx], glht_results$uc$ci_lower),
        coef_ci_upper = c(summary_est$`97.5 %`[time1_active_idx], glht_results$active$ci_upper,
                         summary_est$`97.5 %`[time1_uc_idx], glht_results$uc$ci_upper),
        p_value = c(summary_est$p.value[time1_active_idx], glht_results$active$p_value,
                   summary_est$p.value[time1_uc_idx], glht_results$uc$p_value),
        d = effect_df$d,
        d_ci_lower = effect_df$ci_lower,
        d_ci_upper = effect_df$ci_upper,
        stringsAsFactors = FALSE
      ) %>%
        mutate(across(where(is.numeric), ~round(., 4)))
      
      return(list(
        table = summary_table,
        model = list(
          model = model_check,
          pooled = est,
          glht = glht_results,
          effect_size = effect_df,
          outcome_name = outcome_name,
          model_type = model_label
        )
      ))
    }
  }
  
  # Store results if return_table = TRUE
  if (return_table) {
    models_list <- list()
    tables_list <- list()
  }
  
  # Process all models
  result_imp <- process_model(outcome, "Imputed")
  if (return_table) {
    models_list[["original"]] <- og_summary
    models_list[["imputed"]] <- result_imp$model
    tables_list[["imputed"]] <- result_imp$table
  }
  
  if (worse == "H") {
    result_110 <- process_model(paste0(outcome, "_110"), "Imputed 110%")
    result_120 <- process_model(paste0(outcome, "_120"), "Imputed 120%")
    if (return_table) {
      models_list[["imputed_110"]] <- result_110$model
      tables_list[["imputed_110"]] <- result_110$table
      models_list[["imputed_120"]] <- result_120$model
      tables_list[["imputed_120"]] <- result_120$table
    }
  } else if (worse == "L") {
    result_90 <- process_model(paste0(outcome, "_90"), "Imputed 90%")
    result_80 <- process_model(paste0(outcome, "_80"), "Imputed 80%")
    if (return_table) {
      models_list[["imputed_90"]] <- result_90$model
      tables_list[["imputed_90"]] <- result_90$table
      models_list[["imputed_80"]] <- result_80$model
      tables_list[["imputed_80"]] <- result_80$table
    }
  }
  

  
  if (return_table) {
    return(list(
      summary_table = bind_rows(tables_list),
      models = models_list,
      outcome = outcome
    ))
  }
}
```

## Function: Set the predictor matrix and actual imputation

The df will only consist of the time, week, group, and outcome variables. If needs to add other covariates, update the pred matrix accordingly.

```{r set_the_prediction_matrix}
imputating_missing_long <- function(df_tmp, vars, pred_spec = NULL, restrict_range = c("Y", "N"), range_matrix = NULL){
  set.seed(20250314)
  
  data_long <- df_tmp %>% 
    dplyr::select(id, 
                  time,
                  week,
                  time1,
                  time2,
                  sr_scr_rand_group,
                  all_of(vars),
                  starts_with("missing_"))
  ini <- mice(data_long, maxit=0, print = FALSE)
  pred <- ini$predictorMatrix
  
  pred[,'id'] <- 0
  # pred[, 'time'] <- 0
  pred[, 'week'] <- 0
  pred[, 'time1'] <- 0
  pred[, 'time2'] <- 0
  pred[, 'sr_scr_rand_group'] <- 0
  
  #imputation method: "" = MICE selects the best one automatically; 2l.pan = best for nested continuous data
  meth <- rep("", ncol(data_long))  # Default to MICE choosing method
  names(meth) <- names(data_long)
  if (restrict_range == "Y"){
    post <- make.post(data_long)
  }
  
  for (var in vars) {
    missing_outcome <- paste0('missing_', var)
    pred[, missing_outcome] <- 0
    
    if (is.null(pred_spec)){
    pred[var, ] <- c(-2, 2, 0, 0, 0, 2, rep(2, length(vars)), rep(0, length(vars))) # Update if adding covariates
    } else {
      pred[var, ] <- pred_spec %>% 
        dplyr::select(!!sym(var)) %>% 
        unlist()
    }
    
    if (is.factor(data_long[[var]])) {
      if (length(levels(data_long[[var]])) == 2) {
        meth[var] <- "logreg"  # For binary categorical variables
        } else if (is.ordered(data_long[[var]])) {
          meth[var] <- "polr"    # For ordered categorical variables
          } else {
            meth[var] <- "polyreg" # For unordered categorical variables with >2 levels
            }
      } else if (is.numeric(data_long[[var]])){
      meth[var] <- "2l.pan"
    }
    
    if ((restrict_range == "Y") & (is.numeric(data_long[[var]]))){
      if (var %in% names(range_matrix)) {  # Ensure the variable exists in the range list
        post[var] <- paste0("imp[[j]][, i] <- squeeze(imp[[j]][,i], c(", range_matrix[[var]][1], ", ",
                            range_matrix[[var]][2], "))")
        } else {
          warning(paste("Variable", var, "not found in range_matrix, skipping restriction"))
        }
      }
  }
  
  print(pred)
  print(meth)
  
  if (restrict_range == "Y"){
    imp_tmp <- mice(data_long, 
                    pred = pred, 
                    meth = meth, 
                    maxit = 10, 
                    # print = FALSE, 
                    m = 50,
                    post = post
                    )
  } else {
    imp_tmp <- mice(data_long, 
                    pred = pred, 
                    meth = meth, 
                    maxit = 10, 
                    # print = FALSE, 
                    m = 50
                    )
  }
  
  for (var in vars) {
    missing_outcome <- paste0('missing_', var)
    miss_count <- sum(df_tmp[[missing_outcome]])
    if (miss_count > 0) {  
      print(plot(imp_tmp, var)) 
      print(densityplot(imp_tmp, as.formula(paste("~", var))))
    } else {
      print(paste("Skipping", var, "as it has no missing data and was not imputed"))
    }
  }
  
  return(imp_tmp)
}
```

## Function: Set the predictor matrix and actual imputation

df generated by this code will only consist of group, demographic variables and target variable(s). If needs to add other covariates, update the pred matrix accordingly.

```{r set_the_prediction_matrix}
imputating_missing_wide <- function(df_tmp, vars, pred_spec = NULL, restrict_range = c("Y", "N"), range_matrix = NULL){
  set.seed(20250701) 
  
  df_og <- df_tmp
  
  df_tmp <- df_tmp %>% 
    dplyr::select(id,
                  sr_scr_rand_group, 
                  all_of(vars),
                  starts_with('missing'))
  
  get_base_name <- function(var, value_range_names) {
    matched <- value_range_names[
      vapply(value_range_names, function(x) endsWith(var, x), logical(1))
    ]
    if (length(matched) == 1) matched else NA_character_
  }
  
  ini <- mice(df_tmp, maxit=0, print = FALSE)
  pred <- ini$predictorMatrix
  
  pred[,'id'] <- 0
  pred[, 'sr_scr_rand_group'] <- 0
  pred['id',] <- 0
  
  #imputation method: "" = MICE selects the best one automatically; 2l.pan = best for nested continuous data
  meth <- rep("", ncol(df_tmp))  # Default to MICE choosing method
  names(meth) <- names(df_tmp)
  
  # probably due to skewness, the plasma variable doesn't get a method chosen automatically
  crp_plasma_vars <- grep(
    "^dbs_.*_crp_.*_plasma_.*",
    names(meth),
    value = TRUE
  )
  
  if (length(crp_plasma_vars) > 0) {
    meth[crp_plasma_vars] <- "pmm"
  }
  
  if (restrict_range == "Y"){
    post <- make.post(df_tmp)
  }
  
  for (var in vars) {
    missing_outcome <- paste0('missing_', var)
    pred[, missing_outcome] <- 0
    
      if (is.null(pred_spec)){
    pred[var, ] <- c(0, 1, rep(1, length(vars)), rep(0, length(vars)))
  } else {
    pred[var, ] <- pred_spec %>% 
      dplyr::select(!!sym(var)) %>% 
      unlist()
  }
  
  # Add safety check
  if (sum(pred[var, ]) == 0) {
    warning(paste("Variable", var, "has no predictors!"))
    # Set default predictors
    pred[var, ] <- c(0, 1, rep(1, length(vars)), rep(0, length(vars)))
  }
      
  if (restrict_range == "Y" && is.numeric(df_tmp[[var]])) {
  
    base_var <- get_base_name(var, names(range_matrix))
  
    if (!is.na(base_var)) {
      post[var] <- paste0(
        "imp[[j]][, i] <- squeeze(imp[[j]][, i], c(",
        range_matrix[[base_var]][1], ", ",
        range_matrix[[base_var]][2], "))"
      )
    } else {
      warning(paste("No range found for", var))
    }
  }
  
    }
  
  print(pred)
  
  if (restrict_range == "Y"){
    imp_tmp <- mice(df_tmp, 
                    pred = pred, 
                    maxit = 10, 
                    # print = FALSE, 
                    m = 50,
                    post = post
                    )
  } else {
    imp_tmp <- mice(df_tmp, 
                    pred = pred, 
                    maxit = 10, 
                    # print = FALSE, 
                    m = 50
                    )
  }
  
  has_sufficient_data <- function(imp, var) {
    dat_long <- complete(imp, action = "long", include = TRUE)
    x <- dat_long[[var]]
    x <- x[!is.na(x)]
    length(x) >= 2 && length(unique(x)) > 1
  }
      
  for (var in vars) {
    missing_outcome <- paste0('missing_', var)
    miss_count <- sum(df_og[[missing_outcome]])
  
    if (miss_count > 0 && has_sufficient_data(imp_tmp, var)) {
  
      p1 <- try(plot(imp_tmp, var), silent = TRUE)
      if (!inherits(p1, "try-error")) print(p1)
  
      p2 <- try(densityplot(imp_tmp, as.formula(paste("~", var))), silent = TRUE)
      if (!inherits(p2, "try-error")) print(p2)
  
    } else {
      message("Skipping plots for ", var, 
              ": insufficient variation or no imputation")
    }
  }
    
  
  cat("\n=== MICE Logged Events ===\n")
  print(imp_tmp$loggedEvents)

  return(imp_tmp)
}
```


### Function: scale the imputed values

If higher level = worse, then the scaling should be done with 1.1 and 1.2, while if lower level = worse, then it should be done with 0.8 and 0.9

```{r fx_scale_imputed_wide}
scale_imputed_values_wide <- function(df_tmp, var) {
  worse <- map_worse(var)
  
  var_prefix <- map_prefix(var)
  
  prefixes <- c(paste0(var_prefix, '_base_'),
              paste0(var_prefix, '_pw4_'),
              paste0(var_prefix, '_3mo_'))
  
  for (prefix in prefixes){
    outcome <- paste0(prefix, var)
    missing_var <- paste0("missing_", outcome)
    
    if (!outcome %in% names(df_tmp)) next
    
    if (worse == "H"){
      var110 <- paste0(outcome, "_110")
      var120 <- paste0(outcome, "_120")
      
      df_tmp <- df_tmp %>% 
        mutate(!!sym(var110) := if_else(!!sym(missing_var) == 1, 
                                        !!sym(outcome) * 1.1,
                                        !!sym(outcome)),
               !!sym(var120) := if_else(!!sym(missing_var) == 1, 
                                        !!sym(outcome) * 1.2,
                                        !!sym(outcome))
               )
    } else if (worse == "L"){
      var80 <- paste0(outcome, "_80")
      var90 <- paste0(outcome, "_90")
      
      df_tmp <- df_tmp %>% 
        mutate(!!sym(var80) := if_else(!!sym(missing_var) == 1, 
                                       !!sym(outcome) * 0.8,
                                       !!sym(outcome)),
               !!sym(var90) := if_else(!!sym(missing_var) == 1, 
                                       !!sym(outcome) * 0.9,
                                       !!sym(outcome))
               )
    }
  }
  
  print(tail(df_tmp %>% filter(!!sym(missing_var) == 1)))
  
  return(df_tmp)
}
```

### Function: Simple linear regression model with original data

```{r fx-lm-analysis-wide}
run_lm_og_wide <- function(df_og = df, outcome, phase = '3mo', covar_formula = NULL, return_table = FALSE, exc_outlier = FALSE, ref_level = 'hmp_full'){
  
  df_og <- df_og %>%
    mutate(sr_scr_rand_group = relevel(factor(sr_scr_rand_group), ref = ref_level))
  
  var_prefix <- map_prefix(outcome)
  
  base_var <- paste0(var_prefix, '_base_', outcome)
  outcome_var <- paste0(var_prefix, '_', phase, '_', outcome)
  
  if (exc_outlier == TRUE){ # update df_og if necessary
    if (str_detect(var_prefix, "mib")){
      # only exclude > median + 1.5xIQR
        outcome_med <- median(df_og[[outcome_var]], na.rm = TRUE)
        outcome_IQR_val <- IQR(df_og[[outcome_var]], na.rm = TRUE)
        outcome_upper_bound <- outcome_med + 1.5 * outcome_IQR_val
        base_med <- median(df_og[[base_var]], na.rm = TRUE)
        base_IQR_val <- IQR(df_og[[base_var]], na.rm = TRUE)
        base_upper_bound <- base_med + 1.5 * base_IQR_val

        df_og <- df_og %>%
        mutate(!!sym(outcome_var) := ifelse((!!sym(outcome_var) > outcome_upper_bound), NA, !!sym(outcome_var)),
               !!sym(base_var) := ifelse((!!sym(base_var) > base_upper_bound), NA, !!sym(base_var)))
    } else {
      outcome_mean_y <- mean(df_og[[outcome_var]], na.rm = TRUE)
      outcome_sd_y <- sd(df_og[[outcome_var]], na.rm = TRUE)
      outcome_upper_y <- outcome_mean_y + 3 * outcome_sd_y
      outcome_lower_y <- outcome_mean_y - 3 * outcome_sd_y
      base_mean_y <- mean(df_og[[base_var]], na.rm = TRUE)
      base_sd_y <- sd(df_og[[base_var]], na.rm = TRUE)
      base_upper_y <- base_mean_y + 3 * base_sd_y
      base_lower_y <- base_mean_y - 3 * base_sd_y
      
      df_og <- df_og %>%
        mutate(!!sym(outcome_var) := ifelse((!!sym(outcome_var) > outcome_upper_y | !!sym(outcome_var) < outcome_lower_y), NA, !!sym(outcome_var)),
               !!sym(base_var) := ifelse((!!sym(base_var) > base_upper_y | !!sym(base_var) < base_lower_y), NA, !!sym(base_var)))
    }
  }
  
  cat("\nOutcome:", outcome_var,"(Original, NOT imputed) -----------------------\n")
  og_formula <- as.formula(paste(outcome_var, "~ ", covar_formula, base_var, " + sr_scr_rand_group"))
  cat('\n', as.character(og_formula), '\n\n')
  og_est <- lm(og_formula, data = df_og)
  summary_og_est <- summary(og_est)
  fixed_effects <- summary_og_est$coefficients
  fixed_effects_rounded <- round(fixed_effects, 4)
  
  confint_og_est <- confint(og_est)
  confint_fixed <- confint_og_est[rownames(confint_og_est) %in% rownames(fixed_effects), ]
  confint_fixed_rounded <- round(confint_fixed, 4)
  
  results_table <- cbind(
    Estimate = fixed_effects_rounded[, "Estimate"],
    Std_Error = fixed_effects_rounded[, "Std. Error"],
    t_value = fixed_effects_rounded[, "t value"],
    p_value = fixed_effects_rounded[, "Pr(>|t|)"],
    `CI_2.5%` = confint_fixed_rounded[, 1],
    `CI_97.5%` = confint_fixed_rounded[, 2]
    )
  
  print(results_table)
  
  cat('\n****** Effect Size summary ******\n')
  effect_df <- compute_effect_size_wide(estimation = og_est, 
                                        data = df_og,
                                        base_var = base_var,
                                        outcome_var = outcome_var,
                                        preMI = 'Y',
                                        ref_level = ref_level)
  
  if (return_table == TRUE){
    if(str_detect(outcome_var, '_pw4_')){
      time_desc <- "Week 4 vs. Baseline"
      } else if (str_detect(outcome_var, '_3mo_')){
        time_desc <- "Week 17 vs. Baseline"
      }
    
    if (ref_level == 'hmp_full'){
      group_comparison <- c("Full_vs_Active", "Full_vs_UC")
      group1_term <- "sr_scr_rand_grouphmp_active"
    } else if (ref_level == 'hmp_active'){
      group_comparison <- c("Active_vs_Full", "Active_vs_UC")
      group1_term <- "sr_scr_rand_grouphmp_full"
    }
    
    summary_table <- data.frame(
      outcome = outcome,
      comparison = rep(group_comparison, each = 1),
      timepoint = rep(c(time_desc), 2),
      
      # Coefficients
      coef = c(
        fixed_effects[group1_term, "Estimate"],
        fixed_effects["sr_scr_rand_groupuc", "Estimate"]
        ),
        
      # Coefficient CIs
      coef_ci_lower = c(
        confint_fixed[group1_term, 1],
        confint_fixed["sr_scr_rand_groupuc", 1]
      ),
      coef_ci_upper = c(
        confint_fixed[group1_term, 2],
        confint_fixed["sr_scr_rand_groupuc", 2]
      ),
        
      # P-values
      p_value = c(
        fixed_effects[group1_term, "Pr(>|t|)"],
        fixed_effects["sr_scr_rand_groupuc", "Pr(>|t|)"]
      ),
        
      # Effect sizes
      d = effect_df$d,
      d_ci_lower = effect_df$ci_lower,
      d_ci_upper = effect_df$ci_upper,
      
      stringsAsFactors = FALSE
      )
     
    # Round numeric columns
    summary_table <- summary_table %>%
      mutate(across(where(is.numeric), ~round(., 4)))
    
    return(list(model = og_est, results_summary = summary_table, ci_fixed = confint_fixed))
    }
}
```



### Function: Simple linear regression model with imputed data

There are four models:
- Original model
- Model with imputed data
- Model with imputed data, in which the data is scaled to be 10% worse
- Model with imputed data, in which the data is scaled to be 20% worse

```{r fx-sensitvity-analysis-wide}
run_sensitivity_imputed_wide <- function(df_og, df_imp, outcome, covariates = NULL, return_table = FALSE, exc_outlier = FALSE, ref_level = 'hmp_full'){
  worse <- map_worse(outcome)
  
  df_og <- df_og %>%
    mutate(sr_scr_rand_group = relevel(factor(sr_scr_rand_group), ref = ref_level))
  
  df_imp <- df_imp %>%
    mutate(sr_scr_rand_group = relevel(factor(sr_scr_rand_group), ref = ref_level))
  
  if (!is.null(covariates)){
    covar_formula <- paste0(covariates, collapse=" + ")
    covar_formula <- paste0(covar_formula, "+ ")
  } else {
    covar_formula <- NULL
  }
  
  print(outcome)
  var_prefix <- map_prefix(outcome)
  print(var_prefix)
  
  # Remove redundant prefix from outcome name
  outcome_suffix <- sub(paste0("^", var_prefix, "_"), "", outcome)
  # outcome_suffix <- sub(paste0("_", var_prefix, "_"), "_", outcome_suffix)
  
  base_var <- paste0(var_prefix, '_base_', outcome_suffix)
  outcome_vars <- c(paste0(var_prefix, '_pw4_', outcome_suffix), 
                    paste0(var_prefix, '_3mo_', outcome_suffix))
  
  # Store results if return_table = TRUE
  if (return_table) {
    models_list <- list()
    tables_list <- list()
    ci_fixed_list <- list()
  }
  
  for (outcome_var in outcome_vars){
    # Skip if outcome variable doesn't exist
    if (!outcome_var %in% names(df_og)) {
      cat("Skipping", outcome_var, "- not found in dataframe\n")
      next
    }
    
    if (str_detect(outcome_var, 'pw4')){
      timepoint <- "Week 4 vs. Baseline"
      phase <- "pw4"
      cat("\nTime: Week 4 ===================================================================\n")
    } else if (str_detect(outcome_var, '3mo')){
      timepoint <- "Week 17 vs. Baseline"
      phase <- "3mo"
      cat("\nTime: 3-month follow-up ===================================================================\n")
    }
    
    orig_results <- run_lm_og_wide(df_og = df_og, 
                                   outcome = outcome, 
                                   phase = phase, 
                                   covar_formula = covar_formula, 
                                   exc_outlier = exc_outlier, 
                                   ref_level = ref_level,
                                   return_table = TRUE)
    
    og_summary <- orig_results$results_summary
    og_summary <- og_summary %>% 
      mutate(model_type = "Original") %>% 
      relocate(outcome, model_type)
    
    if (exc_outlier == TRUE){ #exc outlier (>3SDs) if necessary
       df_imp_exc <- df_imp %>%
        group_by(.imp) %>%
        mutate(
          mean_val = mean(!!sym(outcome_var), na.rm = TRUE),
          sd_val = sd(!!sym(outcome_var), na.rm = TRUE),
          threshold_val = (mean_val + 3*sd_val),
          !!sym(outcome_var) := ifelse(
            !!sym(outcome_var) > threshold_val,
            NA,
            !!sym(outcome_var))
          ) %>%
        mutate(
          mean_val_base = mean(!!sym(base_var), na.rm = TRUE),
          sd_val_base = sd(!!sym(base_var), na.rm = TRUE),
          threshold_val_base = (mean_val_base + 3*sd_val_base),
          !!sym(base_var) := ifelse(
            !!sym(base_var) > threshold_val_base,
            NA,
            !!sym(base_var))
          ) %>%
        ungroup() %>%
        dplyr::select(-mean_val, -sd_val, -threshold_val,
                      -mean_val_base, -sd_val_base, -threshold_val_base)

      mids_tmp <- mice::as.mids(df_imp_exc)
      } else {
        mids_tmp <- mice::as.mids(df_imp)
      }
    
    process_model <- function(outcome_name, scaled_perc, model_label) {
      cat("\nOutcome:", outcome_name, "(", model_label, ") -----------------------\n")
      
      outcome_var_scaled <- paste0(outcome_name, scaled_perc)
      base_var_scaled <- paste0(base_var, scaled_perc)
      
      model_check <- with(mids_tmp, 
                          lm(as.formula(paste(outcome_var_scaled, "~ ", covar_formula, base_var_scaled, " + sr_scr_rand_group"))))
      
      est <- pool(model_check)
      summary_est <- summary(est, conf.int = TRUE)
      ci_fixed <- summary_est[, c("2.5 %", "97.5 %")]
rownames(ci_fixed) <- summary_est$term
ci_fixed <- round(ci_fixed, 4)
      summary_est[, 2:ncol(summary_est)] <- round(summary_est[, 2:ncol(summary_est)], 4)
      
      print(summary_est)
      
      effect_df <- compute_effect_size_wide(estimation = est, 
                          data = mids_tmp, 
                          base_var = base_var_scaled, 
                          outcome_var = outcome_var_scaled,
                          preMI = 'N',
                          ref_level = ref_level)
      
      if (return_table) {
        if (ref_level == 'hmp_full'){
          time1_group1_idx <- which(summary_est$term == "sr_scr_rand_grouphmp_active")
          group_comparison = c("Full_vs_Active", "Full_vs_UC")
        } else if (ref_level == 'hmp_active'){
          time1_group1_idx <- which(summary_est$term == "sr_scr_rand_grouphmp_full")
          group_comparison = c("Active_vs_Full", "Active_vs_UC")
        }
        time1_uc_idx <- which(summary_est$term == "sr_scr_rand_groupuc")
        
        summary_table <- data.frame(
          outcome = outcome_name,
          model_type = model_label,
          comparison = rep(group_comparison, each = 1),
          timepoint = rep(timepoint, 2),
          coef = c(summary_est$estimate[time1_group1_idx],
                  summary_est$estimate[time1_uc_idx]),
          coef_ci_lower = c(summary_est$`2.5 %`[time1_group1_idx],
                           summary_est$`2.5 %`[time1_uc_idx]),
          coef_ci_upper = c(summary_est$`97.5 %`[time1_group1_idx],
                           summary_est$`97.5 %`[time1_uc_idx]),
          p_value = c(summary_est$p.value[time1_group1_idx],
                     summary_est$p.value[time1_uc_idx]),
          d = effect_df$d,
          d_ci_lower = effect_df$ci_lower,
          d_ci_upper = effect_df$ci_upper,
          stringsAsFactors = FALSE
        ) %>%
          mutate(across(where(is.numeric), ~round(., 4)))
        
        return(list(
          table = summary_table,
          ci_fixed = ci_fixed, 
          model = list(
            model = model_check,
            pooled = est,
            effect_size = effect_df,
            outcome_name = outcome_name,
            model_type = model_label
          )
        ))
      }
    }
    
    # Process all models
    result_imp <- process_model(outcome_var, "", "Imputed")
    if (return_table && !is.null(result_imp)) {
      og_model_key <- paste0(phase, "_original")
      tables_list[[og_model_key]] <- og_summary
      model_key <- paste0(phase, "_imputed")
      models_list[[model_key]] <- result_imp$model
      tables_list[[model_key]] <- result_imp$table
      ci_fixed_list[[model_key]] <- result_imp$ci_fixed
    }
    
    if (worse == "H") {
      result_110 <- process_model(outcome_var, "_110", "Imputed 110%")
      result_120 <- process_model(outcome_var, "_120", "Imputed 120%")
      if (return_table) {
        if (!is.null(result_110)) {
          model_key <- paste0(phase, "_imputed_110")
          models_list[[model_key]] <- result_110$model
          tables_list[[model_key]] <- result_110$table
          ci_fixed_list[[model_key]] <- result_imp$ci_fixed
        }
        if (!is.null(result_120)) {
          model_key <- paste0(phase, "_imputed_120")
          models_list[[model_key]] <- result_120$model
          tables_list[[model_key]] <- result_120$table
          ci_fixed_list[[model_key]] <- result_imp$ci_fixed
        }
      }
    } else if (worse == "L") {
      result_90 <- process_model(outcome_var, "_90", "Imputed 90%")
      result_80 <- process_model(outcome_var, "_80", "Imputed 80%")
      if (return_table) {
        if (!is.null(result_90)) {
          model_key <- paste0(phase, "_imputed_90")
          models_list[[model_key]] <- result_90$model
          tables_list[[model_key]] <- result_90$table
          ci_fixed_list[[model_key]] <- result_imp$ci_fixed
        }
        if (!is.null(result_80)) {
          model_key <- paste0(phase, "_imputed_80")
          models_list[[model_key]] <- result_80$model
          tables_list[[model_key]] <- result_80$table
          ci_fixed_list[[model_key]] <- result_imp$ci_fixed
        }
      }
    }
  }
  
  if (return_table) {
    if (length(tables_list) == 0) {
      cat("Warning: No valid models were created for", outcome, "\n")
      return(list(
        summary_table = data.frame(),
        models = list(),
        outcome = outcome
      ))
    }
    
    return(list(
      summary_table = bind_rows(tables_list),
      ci_fixed = ci_fixed_list,
      models = models_list,
      outcome = outcome
    ))
  }
}
```


### Function: Simple linear regression model with original data and moderation

```{r fx-lm-analysis-wide-moderation}
run_lm_og_wide_moderation <- function(df_og = df, outcome, phase = '3mo', covar_formula = NULL, return_table = FALSE, moderator_var){
    var_prefix <- map_prefix(outcome)
  
    base_var <- paste0(var_prefix, '_base_', outcome)
    outcome_var <- paste0(var_prefix, '_', phase, '_', outcome)
    
    activeFull <- "sr_scr_rand_grouphmp_active"
    ucFull <- "sr_scr_rand_groupuc"
    activeFull_moderation <- paste0(activeFull, ":", moderator_var)
    ucFull_moderation <- paste0(ucFull, ":", moderator_var)
    
    cat("\nOutcome:", outcome_var,"(Original, NOT imputed) -----------------------\n")
    cat("\nModerator:", moderator_var,"-----------------------\n")

    og_formula <- as.formula(paste(outcome_var, "~ ", covar_formula, base_var, " + sr_scr_rand_group*", moderator_var))

    cat('\n', as.character(og_formula), '\n\n')
    og_est <- lm(og_formula, data = df_og)
    summary_og_est <- summary(og_est)
    fixed_effects <- summary_og_est$coefficients
    fixed_effects_rounded <- round(fixed_effects, 4)
    
    confint_og_est <- confint(og_est)
    confint_fixed <- confint_og_est[rownames(confint_og_est) %in% rownames(fixed_effects), ]
    confint_fixed_rounded <- round(confint_fixed, 4)
    
    results_table <- cbind(
    Estimate = fixed_effects_rounded[, "Estimate"],
    Std_Error = fixed_effects_rounded[, "Std. Error"],
    t_value = fixed_effects_rounded[, "t value"],
    p_value = fixed_effects_rounded[, "Pr(>|t|)"],
    `CI_2.5%` = confint_fixed_rounded[, 1],
    `CI_97.5%` = confint_fixed_rounded[, 2]
    )
    
    print(results_table)
    
    cat('\n****** Effect Size summary ******\n')
    effect_df <- compute_effect_size_wide_moderation(estimation = og_est, 
                                          data = df_og,
                                          base_var = base_var, 
                                          outcome_var = outcome_var,
                                          preMI = 'Y',
                                          moderator_var = moderator_var)
    
    if (return_table == TRUE){
      if(str_detect(outcome_var, '_pw4_')){
        time_desc <- "Week 4 vs. Baseline"
      } else if (str_detect(outcome_var, '_3mo_')){
        time_desc <- "Week 17 vs. Baseline"
      }
      summary_table <- data.frame(
        outcome = outcome,
        comparison = rep(c("Full_vs_Active", "Full_vs_UC"), each = 3),
        timepoint = rep(c(time_desc), 2),
        effect = effect_df$effect,
        
        # Coefficients
        coef = c(
          fixed_effects[activeFull, "Estimate"],
          fixed_effects[activeFull_moderation, "Estimate"],
          NA,
          fixed_effects[ucFull, "Estimate"],
          fixed_effects[ucFull_moderation, "Estimate"],
          NA
        ),
        
        # Coefficient CIs
        coef_ci_lower = c(
          confint_fixed[activeFull, 1],
          confint_fixed[activeFull_moderation, 1],
          NA,
          confint_fixed[ucFull, 1],
          confint_fixed[ucFull_moderation, 1],
          NA
        ),
        coef_ci_upper = c(
          confint_fixed[activeFull, 2],
          confint_fixed[activeFull_moderation, 2],
          NA,
          confint_fixed[ucFull, 2],
          confint_fixed[ucFull_moderation, 2],
          NA
        ),
        
        # P-values
        p_value = c(
          fixed_effects[activeFull, "Pr(>|t|)"],
          fixed_effects[activeFull_moderation, "Pr(>|t|)"],
          NA,
          fixed_effects[ucFull, "Pr(>|t|)"],
          fixed_effects[ucFull_moderation, "Pr(>|t|)"],   
          NA
        ),
        
        # Effect sizes
        d = effect_df$d,
        d_ci_lower = effect_df$ci_lower,
        d_ci_upper = effect_df$ci_upper,
        
        stringsAsFactors = FALSE
      )
     
       # Round numeric columns
    summary_table <- summary_table %>%
      mutate(across(where(is.numeric), ~round(., 4)))
    
    return(list(model = og_est, results_summary = summary_table, ci_fixed = confint_fixed))

    }
}
```


### Function: Effect size computation for wide format data

```{r fx_effect_size_imputed_wide}
compute_effect_size_wide <- function(estimation, data, base_var, outcome_var, preMI = c('Y', 'N'), ref_level = 'hmp_full'){
  preMI <- match.arg(preMI)
  
  if (ref_level == 'hmp_full'){
    group1_term <- "sr_scr_rand_grouphmp_active"
    group_comparison <- c("full_vs_active", 
                          "full_vs_uc")
  } else if (ref_level == 'hmp_active'){
    group1_term <- "sr_scr_rand_grouphmp_full"
    group_comparison <- c("active_vs_full", 
                          "active_vs_uc")
  }
  
  if (preMI == 'Y') {
    if (inherits(estimation, "lm") && !inherits(estimation, "vglm")) {
      sum_coef <- summary(estimation)$coefficients
      
      est_activeFull <- sum_coef[group1_term, "Estimate"]
      est_ucGroup1 <- sum_coef["sr_scr_rand_groupuc", "Estimate"]
      
    } else {
      # everything else: vglm (Tobit), etc.
      coefs <- coef(estimation)
      if (is.matrix(coefs)) coefs <- coefs[, 1]
      
      est_activeFull <- coefs[group1_term]
      est_ucGroup1 <- coefs["sr_scr_rand_groupuc"]
    }

  } else if (preMI == 'N') {
  est_activeFull <- summary(estimation) %>% 
    filter(term == group1_term) %>% 
    pull(estimate)
  est_ucGroup1 <- summary(estimation) %>% 
    filter(term == "sr_scr_rand_groupuc") %>% 
    pull(estimate)
  }
  
  if (inherits(data, "mids")) {
    # Use the first imputed dataset for calculating sample sizes and SDs
    data <- mice::complete(data, action = 1)
  } 
  
  #stat from original df~~~~~~~~~~~~~
  full_n <- data %>% filter(sr_scr_rand_group == 'hmp_full' & 
                            !is.na(!!sym(base_var)) &
                            !is.na(!!sym(outcome_var))) %>% nrow()
  active_n <- data %>% filter(sr_scr_rand_group == 'hmp_active' & 
                            !is.na(!!sym(base_var)) &
                            !is.na(!!sym(outcome_var))) %>% nrow()
  uc_n <- data %>% filter(sr_scr_rand_group == 'uc' & 
                            !is.na(!!sym(base_var)) &
                            !is.na(!!sym(outcome_var))) %>% nrow()
  sd_base_full <- sd((data %>%
                        filter(sr_scr_rand_group %in% c('hmp_full')) %>%
                        dplyr::select(!!sym(base_var)) %>% 
                        unlist()),
                     na.rm = TRUE)
  sd_base_active <- sd((data %>%
                        filter(sr_scr_rand_group %in% c('hmp_active')) %>%
                        dplyr::select(!!sym(base_var)) %>% 
                        unlist()),
                     na.rm = TRUE)
  sd_base_uc <- sd((data %>% 
                          filter(sr_scr_rand_group %in% c('uc')) %>%
                          dplyr::select(!!sym(base_var)) %>% 
                          unlist()),
                       na.rm = TRUE)
  sd_base_activeFull <- sqrt(((full_n - 1)*sd_base_full^2 + (active_n - 1)*sd_base_active^2) / (full_n + active_n - 2))
  sd_base_ucFull <- sqrt(((full_n - 1)*sd_base_full^2 + (uc_n - 1)*sd_base_uc^2) / (full_n + uc_n - 2))
  sd_base_ucActive <- sqrt(((active_n - 1)*sd_base_active^2 + (uc_n - 1)*sd_base_uc^2) / (active_n + uc_n - 2))
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~
  
  smd_activeFull <- est_activeFull / sd_base_activeFull
  ci_activeFull <- ci.smd(smd = smd_activeFull, n.1 = full_n, n.2 = active_n)
  
  cat('\nIntervention vs Active Ctrl******\n')
  cat('time 1 x group Estimate:', est_activeFull, '\n')
  cat('Intervention sample size:', full_n, '\n')
  cat('Active Ctrl sample size:', active_n, '\n')
  cat('Intervention baseline SD:', round(sd_base_full, 4), '\n')
  cat('Active Ctrl baseline SD:', round(sd_base_active, 4), '\n')
  cat('Pooled SD:', round(sd_base_activeFull, 4), '\n')
  cat('Effect size:\n', 
      est_activeFull, ' /', sd_base_activeFull, '=', smd_activeFull, '\n')
  cat('95% CI:', round(ci_activeFull$Lower, 4), ',', round(ci_activeFull$Upper, 4), '\n')
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~
  if (ref_level == 'hmp_full'){
    smd_ucGroup1 <- est_ucGroup1 / sd_base_ucFull
    ci_ucGroup1 <- ci.smd(smd = smd_ucGroup1, n.1 = full_n, n.2 = uc_n)
    
    cat('\nIntervention vs Usual Care******\n')
    cat('time 1 x group Estimate:', est_ucGroup1, '\n')
    cat('Intervention sample size:', full_n, '\n')
    cat('Usual Care sample size:', uc_n, '\n')
    cat('Intervention baseline SD:', round(sd_base_full, 4), '\n')
    cat('Usual Care baseline SD:', round(sd_base_uc, 4), '\n')
    cat('Pooled SD:', round(sd_base_ucFull, 4), '\n')
    cat('Effect size:\n', 
        est_ucGroup1, ' /', sd_base_ucFull, '=', smd_ucGroup1, '\n')
    cat('95% CI:', round(ci_ucGroup1$Lower, 4), ',', round(ci_ucGroup1$Upper, 4), '\n')
  } else if (ref_level == 'hmp_active'){
    smd_ucGroup1 <- est_ucGroup1 / sd_base_ucActive
    ci_ucGroup1 <- ci.smd(smd = smd_ucGroup1, n.1 = active_n, n.2 = uc_n)
    
    cat('\nActive Ctrl vs Usual Care******\n')
    cat('time 1 x group Estimate:', est_ucGroup1, '\n')
    cat('Active Ctrl sample size:', active_n, '\n')
    cat('Usual Care sample size:', uc_n, '\n')
    cat('Active Ctrl baseline SD:', round(sd_base_active, 4), '\n')
    cat('Usual Care baseline SD:', round(sd_base_uc, 4), '\n')
    cat('Pooled SD:', round(sd_base_ucActive, 4), '\n')
    cat('Effect size:\n', 
        est_ucGroup1, ' /', sd_base_ucActive, '=', smd_ucGroup1, '\n')
    cat('95% CI:', round(ci_ucGroup1$Lower, 4), ',', round(ci_ucGroup1$Upper, 4), '\n')
  }
  
  
  effect_df <- data.frame(
    group = group_comparison,
    effect_type = c(
      "T1",
      "T1"
    ),
    d = c(
      smd_activeFull,
      smd_ucGroup1
    ),
    ci_lower = c(
      ci_activeFull$Lower,
      ci_ucGroup1$Lower
    ),
    ci_upper = c(
      ci_activeFull$Upper,
      ci_ucGroup1$Upper
    )
  )
  
  return(effect_df)
}
```


### Function: Effect size computation with moderation for wide format data

```{r fx_effect_size_imputed_wide_moderation}
compute_effect_size_wide_moderation <- function(estimation, data, base_var, outcome_var, preMI = c('Y', 'N'), moderator_var){
  preMI <- match.arg(preMI)
  
  active_moderation <- paste0("sr_scr_rand_grouphmp_active:", moderator_var)
  uc_moderation <- paste0("sr_scr_rand_groupuc:", moderator_var)  
    
  if (preMI == 'Y') {
    sum_coef <- summary(estimation)$coefficients
    
    est_activeFull <- sum_coef["sr_scr_rand_grouphmp_active", "Estimate"]
    est_activeFull_moderation <- sum_coef[active_moderation, "Estimate"]
    est_ucFull <- sum_coef["sr_scr_rand_groupuc", "Estimate"]
    est_ucFull_moderation <- sum_coef[uc_moderation, "Estimate"]
  } else if (preMI == 'N') {
  est_activeFull <- summary(estimation) %>% 
    filter(term == "sr_scr_rand_grouphmp_active") %>% 
    pull(estimate)
  est_activeFull_moderation <- summary(estimation) %>% 
    filter(term == active_moderation) %>% 
    pull(estimate)
  est_ucFull <- summary(estimation) %>% 
    filter(term == "sr_scr_rand_groupuc") %>% 
    pull(estimate)
  est_ucFull_moderation <- summary(estimation) %>% 
    filter(term == uc_moderation) %>% 
    pull(estimate)
  }
  
  if (inherits(data, "mids")) {
    # Use the first imputed dataset for calculating sample sizes and SDs
    data <- mice::complete(data, action = 1)
  } 
  
  #stat from original df~~~~~~~~~~~~~
  full_n <- data %>% filter(sr_scr_rand_group == 'hmp_full' & 
                            !is.na(!!sym(base_var)) &
                            !is.na(!!sym(outcome_var))) %>% nrow()
  active_n <- data %>% filter(sr_scr_rand_group == 'hmp_active' & 
                            !is.na(!!sym(base_var)) &
                            !is.na(!!sym(outcome_var))) %>% nrow()
  uc_n <- data %>% filter(sr_scr_rand_group == 'uc' & 
                            !is.na(!!sym(base_var)) &
                            !is.na(!!sym(outcome_var))) %>% nrow()

  sd_base_full <- sd((data %>%
                        filter(sr_scr_rand_group %in% c('hmp_full')) %>%
                        dplyr::select(!!sym(base_var)) %>% 
                        unlist()),
                     na.rm = TRUE)
  sd_base_active <- sd((data %>%
                        filter(sr_scr_rand_group %in% c('hmp_active')) %>%
                        dplyr::select(!!sym(base_var)) %>% 
                        unlist()),
                     na.rm = TRUE)
  sd_base_uc <- sd((data %>% 
                          filter(sr_scr_rand_group %in% c('uc')) %>%
                          dplyr::select(!!sym(base_var)) %>% 
                          unlist()),
                       na.rm = TRUE)
  sd_base_activeFull <- sqrt(((full_n - 1)*sd_base_full^2 + (active_n - 1)*sd_base_active^2) / (full_n + active_n - 2))
  sd_base_ucFull <- sqrt(((full_n - 1)*sd_base_full^2 + (uc_n - 1)*sd_base_uc^2) / (full_n + uc_n - 2))
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~
  
  # mean of moderator
  moderator_mean <- mean(data[[moderator_var]], na.rm = TRUE)
  
  smd_activeFull <- est_activeFull / sd_base_activeFull
  ci_activeFull <- ci.smd(smd = smd_activeFull, n.1 = full_n, n.2 = active_n)
  
  smd_activeFull_moderation <- (est_activeFull_moderation*moderator_mean) / sd_base_activeFull
  ci_activeFull_moderation <- ci.smd(smd = smd_activeFull_moderation, n.1 = full_n, n.2 = active_n)
  
  smd_activeFull_full <- (est_activeFull + (est_activeFull_moderation*moderator_mean)) / sd_base_activeFull
  ci_activeFull_full <- ci.smd(smd = smd_activeFull_full, n.1 = full_n, n.2 = active_n)
  
  cat('\nMean of', moderator_var ,'******\n')
  cat(moderator_mean, '\n')
  
  cat('\nIntervention vs Active Ctrl******\n')
  cat('Group Estimate:', est_activeFull, '\n')
  cat('Group *', moderator_var, 'Estimate:', est_activeFull_moderation, '\n')
  cat('Intervention sample size:', full_n, '\n')
  cat('Active Ctrl sample size:', active_n, '\n')
  cat('Group effect size:\n', 
      est_activeFull, ' /', sd_base_activeFull, '=', smd_activeFull, '\n')
  cat('95% CI:', round(ci_activeFull$Lower, 4), ',', round(ci_activeFull$Upper, 4), '\n')
  
  cat('Interaction effect size:\n', 
      est_activeFull_moderation, '*', moderator_mean, '/', sd_base_activeFull, '=', smd_activeFull_moderation, '\n')
  cat('95% CI:', round(ci_activeFull_moderation$Lower, 4), ',', round(ci_activeFull_moderation$Upper, 4), '\n')
  
  cat('Full treatment effect size:\n', 
      '(', est_activeFull, '+ (', est_activeFull_moderation, '*', moderator_mean, '))/', sd_base_activeFull, '=', smd_activeFull_full, '\n')
  cat('95% CI:', round(ci_activeFull_full$Lower, 4), ',', round(ci_activeFull_full$Upper, 4), '\n')
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~
  
  smd_ucFull <- est_ucFull / sd_base_ucFull
  ci_ucFull <- ci.smd(smd = smd_ucFull, n.1 = full_n, n.2 = uc_n)
  
  smd_ucFull_moderation <- (est_ucFull_moderation*moderator_mean) / sd_base_ucFull
  ci_ucFull_moderation <- ci.smd(smd = smd_ucFull_moderation, n.1 = full_n, n.2 = uc_n)
  
  smd_ucFull_full <- (est_ucFull + (est_ucFull_moderation*moderator_mean)) / sd_base_ucFull
  ci_ucFull_full <- ci.smd(smd = smd_ucFull_full, n.1 = full_n, n.2 = uc_n)
  
  cat('\nIntervention vs Usual Care******\n')
  cat('Group Estimate:', est_ucFull, '\n')
  cat('Group *', moderator_var, 'Estimate:', est_ucFull_moderation, '\n')
  cat('Intervention sample size:', full_n, '\n')
  cat('Usual Care sample size:', uc_n, '\n')
  cat('Group effect size:\n', 
      est_ucFull, ' /', sd_base_ucFull, '=', smd_ucFull, '\n')
  cat('95% CI:', round(ci_ucFull$Lower, 4), ',', round(ci_ucFull$Upper, 4), '\n')
  
  cat('Interaction effect size:\n', 
      est_ucFull_moderation, '*', moderator_mean, '/', sd_base_ucFull, '=', smd_ucFull_moderation, '\n')
  cat('95% CI:', round(ci_ucFull_moderation$Lower, 4), ',', round(ci_ucFull_moderation$Upper, 4), '\n')
  
  cat('Full treatment effect size:\n', 
      '(', est_ucFull, '+ (', est_ucFull_moderation, '*', moderator_mean, '))/', sd_base_ucFull, '=', smd_ucFull_full, '\n')
  cat('95% CI:', round(ci_ucFull_full$Lower, 4), ',', round(ci_ucFull_full$Upper, 4), '\n')
  
  effect_df <- data.frame(
    group = c(
      "full_vs_active", "full_vs_active", "full_vs_active", 
      "full_vs_uc", "full_vs_uc", "full_vs_uc"
    ),
    effect_type = c(
      "T1", "T1", "T1",
      "T1", "T1","T1"
    ),
    effect = c(
      "Group", "Interaction", "Full effect",
      "Group", "Interaction", "Full effect"
      ),
    d = c(
      smd_activeFull, smd_activeFull_moderation, smd_activeFull_full,
      smd_ucFull, smd_ucFull_moderation, smd_ucFull_full
    ),
    ci_lower = c(
      ci_activeFull$Lower, ci_activeFull_moderation$Lower, ci_activeFull_full$Lower,
      ci_ucFull$Lower, ci_ucFull_moderation$Lower, ci_ucFull_full$Lower
    ),
    ci_upper = c(
      ci_activeFull$Upper, ci_activeFull_moderation$Upper, ci_activeFull_full$Upper,
      ci_ucFull$Upper, ci_ucFull_moderation$Upper, ci_ucFull_full$Upper
    )
  )
  
  return(effect_df)
}
```


## DBS functions
### For tables
```{r dbs_tables}

dbs_results <- function(est, target_var, smd_results, model_name, ref_level = 'hmp_full'){

# extract coef and CI
# lm: uses confint(), which uses t — correct for OLS
# vglm: uses 1.96 * se — correct for MLE/Tobit
  if (inherits(est, "lm") && !inherits(est, "vglm")) {
    summary_est <- summary(est)
    fixed_effects <- summary_est$coefficients
    
    confint_est <- confint(est)
    confint_fixed <- confint_est[rownames(confint_est) %in% rownames(fixed_effects), ]
  
  } else {
    # vglm / Tobit
    summary_est <- summary(est)
    coef_mat <- coef(summary_est)  # matrix with Estimate, Std. Error, z, p
    
    # Keep only the mean equation rows (drop the log(scale) row for Tobit)
    fixed_effects <- coef_mat[!grepl("^log\\(scale\\)", rownames(coef_mat)), , drop = FALSE]
    
    z <- qnorm(0.975)
    se <- fixed_effects[, "Std. Error"]
    estimate <- fixed_effects[, "Estimate"]
        
    # p-value column is actually Pr(>|z|) 
    # Rename p-value column to match lm's "Pr(>|t|)" so table indexing works uniformly
    p_col <- grep("^Pr", colnames(fixed_effects), value = TRUE)
    colnames(fixed_effects)[colnames(fixed_effects) == p_col] <- "Pr(>|t|)"
    
    confint_fixed <- cbind(
      "2.5 %"  = estimate - z * se,
      "97.5 %" = estimate + z * se
    )
  }
    
    fixed_effects_rounded <- round(fixed_effects, 4)
    confint_fixed_rounded <- round(confint_fixed, 4)

    time_desc <- "Week 17 vs. Baseline"
    
      
  if (ref_level == 'hmp_full'){
    group1_term <- "sr_scr_rand_grouphmp_active"
    group_comparison <- c("Full_vs_Active", 
                          "Full_vs_UC")
  } else if (ref_level == 'hmp_active'){
    group1_term <- "sr_scr_rand_grouphmp_full"
    group_comparison <- c("Active_vs_Full", 
                          "Active_vs_UC")
  }
    
    summary_table <- data.frame(
        outcome = target_var,
        comparison = rep(group_comparison, each = 1),
        timepoint = rep(c(time_desc), 2),
        model = model_name,
        
        # Coefficients
        coef = c(
          fixed_effects[group1_term, "Estimate"],
          fixed_effects["sr_scr_rand_groupuc", "Estimate"]
        ),
        
        # Coefficient CIs
        coef_ci_lower = c(
          confint_fixed[group1_term, 1],
          confint_fixed["sr_scr_rand_groupuc", 1]
        ),
        coef_ci_upper = c(
          confint_fixed[group1_term, 2],
          confint_fixed["sr_scr_rand_groupuc", 2]
        ),
        
        # P-values
        p_value = c(
          fixed_effects[group1_term, "Pr(>|t|)"],
          fixed_effects["sr_scr_rand_groupuc", "Pr(>|t|)"]
        ),
        
        # Effect sizes
        d = smd_results$d,
        d_ci_lower = smd_results$ci_lower,
        d_ci_upper = smd_results$ci_upper,
        
        stringsAsFactors = FALSE
      )
     
       # Round numeric columns
    summary_table <- summary_table %>%
      mutate(across(where(is.numeric), ~round(., 4)))
    
    return(list(
      summary_table = summary_table,
      ci_fixed = confint_fixed_rounded
    ))
}
```

### Function: Tobit effect size
```{r tobit_effect_size}
compute_effect_size_tobit <- function(fit, data, base_var, outcome_var, 
                                       ref_level = 'hmp_full', left_val, B = 2000, seed = 20260521) {
  
  if (ref_level == 'hmp_full') {
    group1_term    <- "sr_scr_rand_grouphmp_active"
    group_comparison <- c("full_vs_active", "full_vs_uc")
  } else if (ref_level == 'hmp_active') {
    group1_term    <- "sr_scr_rand_grouphmp_full"
    group_comparison <- c("active_vs_full", "active_vs_uc")
  }
  
  beta_names   <- c(group1_term, "sr_scr_rand_groupuc")
  log_sig_name <- "(Intercept):2"
  
  # ── Point estimates ──────────────────────────────────────────────────────────
  cf        <- coef(fit)
  sigma_hat <- exp(cf[log_sig_name])
  d_hat     <- cf[beta_names] / sigma_hat
  
  # ── Bootstrap CIs ────────────────────────────────────────────────────────────
  boot_one <- function(i) {
    idx   <- sample.int(nrow(data), replace = TRUE)
    fit_b <- tryCatch(
      vglm(formula(fit),
           family = tobit(Lower = left_val),
           data   = data[idx, ]),
      error = function(e) NULL)
    if (is.null(fit_b)) return(rep(NA_real_, length(beta_names)))
    
    cf_b  <- coef(fit_b)
    
    if (!all(c(beta_names, log_sig_name) %in% names(cf_b))) return(rep(NA_real_, length(beta_names)))
    
    sig_b <- exp(cf_b[log_sig_name])
    
    unname(cf_b[beta_names] / sig_b)
  }
  
  set.seed(seed)
  boot_mat <- t(sapply(seq_len(B), boot_one))
  
  prop_failed <- mean(is.na(boot_mat[, 1]))
  if (prop_failed > 0.05) warning(sprintf("%.1f%% of bootstrap resamples failed to converge", prop_failed * 100))
  
  ci_lower <- apply(boot_mat, 2, quantile, probs = 0.025, na.rm = TRUE)
  ci_upper <- apply(boot_mat, 2, quantile, probs = 0.975, na.rm = TRUE)
  
  # ── Diagnostic printout ─────────────
  cat('\nTobit SMD (sigma from model)\n')
  cat('sigma_hat:', round(sigma_hat, 4), '\n')
  cat(sprintf('%s  d = %.4f  95%% CI [%.4f, %.4f]\n',
              group_comparison[1], d_hat[1], ci_lower[1], ci_upper[1]))
  cat(sprintf('%s  d = %.4f  95%% CI [%.4f, %.4f]\n',
              group_comparison[2], d_hat[2], ci_lower[2], ci_upper[2]))
  cat(sprintf('Bootstrap convergence failures: %.1f%%\n', prop_failed * 100))
  
  # ── Return in same shape as compute_effect_size_wide ─────────────────────────
  effect_df <- data.frame(
    group      = group_comparison,
    effect_type = "T1",
    d          = unname(d_hat),
    ci_lower   = unname(ci_lower),
    ci_upper   = unname(ci_upper)
  )
  
  return(effect_df)
}
```


### Function: Tobit regression for DBS data
```{r tobit_comparison}
tobit_comparison_eh <- function(df_tmp, target_var, exc_outlier = FALSE, ref_level = 'hmp_full'){
  model_list <- list()
  df_tmp <- df_tmp %>%
    mutate(sr_scr_rand_group = relevel(factor(sr_scr_rand_group), ref = ref_level))
  
  # create baseline appending dbs_base to target
  base_var <- paste0('dbs_base_', target_var)
  # create followup (3mo) appending dbs_3mo to target 
  fu_var <- paste0('dbs_3mo_', target_var)
  
   # --- LLD / left-censoring for Tobit ---
  left_val <- ifelse(
    str_detect(target_var, "il6|il10"), 0.4,
    ifelse(str_detect(target_var, "tnf"), 0.9, NA_real_))  # CRP or other uncensored
  
  if (exc_outlier == TRUE){
    mean_y <- mean(df_tmp[[fu_var]], na.rm = TRUE)
    sd_y <- sd(df_tmp[[fu_var]], na.rm = TRUE)
    threshold_upper_y <- mean_y + 3 * sd_y
    threshold_lower_y <- mean_y - 3 * sd_y
    
    mean_base <- mean(df_tmp[[base_var]], na.rm = TRUE)
    sd_base <- sd(df_tmp[[base_var]], na.rm = TRUE)
    threshold_upper_base <- mean_base + 3 * sd_base
    threshold_lower_base <- mean_base - 3 * sd_base
    
    df_tmp <- df_tmp %>%
      mutate(!!sym(fu_var) := ifelse(
               !!sym(fu_var) > threshold_upper_y | !!sym(fu_var) < threshold_lower_y,
               NA, !!sym(fu_var)),
             !!sym(base_var) := ifelse(
               !!sym(base_var) > threshold_upper_base | !!sym(base_var) < threshold_lower_base,
               NA, !!sym(base_var)))
    }


  
  # filters CRP>10
  df_tmp_exc2 <- df_tmp |> 
    filter(dbs_base_crp_plasma_mean_AB <= 10 &
           dbs_3mo_crp_plasma_mean_AB <= 10)
  threshold_string <- "CRP > 10 at baseline or 3mo"

# residual plotting 
plot_residuals <- function(model, df, model_name, target_var) {
  # identify which vars the model used 
  model_vars <- all.vars(formula(model)) 
  
  # find rows w complete data on those vars 
  #used_rows <- setdiff(seq_len(nrow(df)), model@na.action) 
  rows_used <- which(complete.cases(df[, model_vars, drop = FALSE]))
  
  # extract residuals aligned with rows_used 
  res <- residuals(model, type = "response")
  #res <- res[seq_along(used_rows)] # keep only real obsv 
  
  # create full-length residual vector 
  res_full <- rep(NA_real_, nrow(df))
  res_full[rows_used] <- res 

  # build plotting df              
  residuals_df <- data.frame(Residuals = res_full, Group = df$sr_scr_rand_group)
  
  # plot
  ggplot(residuals_df, aes(x = as.factor(Group), y = Residuals, fill = as.factor(Group))) +
    geom_violin(trim = FALSE, alpha = 0.4, color = "black") +  
    geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +  
    geom_jitter(width = 0.1, size = 1, alpha = 0.6, color = "black") + 
    labs(
      title = paste("Residuals Violin Plot:", target_var, "-", model_name), 
      x = "Randomization Group", 
      y = "Residuals"
    ) +
    theme_minimal() +
    theme(
      legend.position = "none",
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text = element_text(size = 12),
      axis.title = element_text(size = 14)
    )
}

  # Formula for ALL models: 
  formula_full <- as.formula(paste0(fu_var, "~ sr_base_curhealth_2infection + sr_3mo_curhealth_2infection + ", base_var, " + sr_scr_rand_group"))
  ## formula_simple is for analyses that exclude anyone with 2+ infection sx 
  formula_simple <- as.formula(paste0(fu_var, "~", base_var, " + sr_scr_rand_group")) 
  
  # run tobit 
  run_model <- function(formula, data, model_name) {
    if(!is.na(left_val)) {
      model <- vglm(formula, tobit(Lower = left_val), data = data)
    } else {
      model <- lm(formula, data = data)
    }
    return(model)
  }
  
  cat("Running tobit regression model for:", target_var, "========================================\n")
  # Model 1: all samples, infectious symptoms as covariate
  cat("Model 1: Run all samples with infectious symptoms as covariate------------\n")
  
  print(describeBy(df_tmp %>% dplyr::select(contains(target_var)), group = df_tmp$sr_scr_rand_group, digits = 3))
  
  model1 <- run_model(formula_full, df_tmp, "Model 1")
  
  cat('\n~~~~~~~~~~~~~~~~\nModel 1: ', 'dbs_3mo_', target_var, ' ~ sr_base_curhealth_2infection + sr_3mo_curhealth_2infection + dbs_base_', target_var, ' + sr_scr_rand_group', '\n~~~~~~~~~~~~~~~~\n')
  
  print(summary(model1))
  print(plot_residuals(model1, df_tmp, "Model 1", target_var))
  
  lm_model_1_d <- compute_effect_size_tobit(fit = model1, 
                                            data = df_tmp %>% filter(!is.na(!!sym(base_var)) & !is.na(!!sym(fu_var))), 
                                            base_var = base_var,
                                            outcome_var = fu_var,
                                            left_val = left_val,
                                            ref_level = ref_level)
  lm_model1_results_all <- dbs_results(est = model1,
                                   target_var = target_var,
                                   smd_results =lm_model_1_d,
                                   model_name = "Original",
                                   ref_level = ref_level)
  lm_model1_results <- lm_model1_results_all$summary_table
  lm_model1_ci <- lm_model1_results_all$ci_fixed
  model_list[["Original"]] <- model1
  
  
  # Model 2: excluding CRP >10, infectious symptoms as covariate 
  cat("\nModel 2: Re-run excluding CRP > 10 with infectious symptoms as covariate------------\n")
  print(describeBy(df_tmp_exc2 %>% 
                     dplyr::select(contains(target_var)), group = df_tmp_exc2$sr_scr_rand_group, digits = 3))
  
  model2 <- run_model(formula_full, df_tmp_exc2, "Model 2") 
  
  cat('\n~~~~~~~~~~~~~~~~\nModel 2: ', 'dbs_3mo_', target_var, ' ~ sr_base_curhealth_2infection + sr_3mo_curhealth_2infection + dbs_base_', target_var, ' + sr_scr_rand_group', '\n~~~~~~~~~~~~~~~~\n')

  print(summary(model2))
  print(plot_residuals(model2, df_tmp, "Model 2", target_var))
  
    
  lm_model_2_d <- compute_effect_size_tobit(fit = model2,
                                            data = df_tmp_exc2 %>% filter(!is.na(!!sym(base_var)) &
                                                                            !is.na(!!sym(fu_var))),
                                            base_var = base_var,
                                            outcome_var = fu_var,
                                            left_val = left_val,
                                            ref_level = ref_level)
  lm_model2_results_all <- dbs_results(est = model2,
                                   target_var = target_var,
                                    smd_results =lm_model_2_d,
                                    model_name = "Excluding CRP>10 with infectious symptoms as covariate",
                                   ref_level = ref_level)
  lm_model2_results <- lm_model2_results_all$summary_table
  lm_model2_ci <- lm_model2_results_all$ci_fixed
  model_list[["crp"]] <- model2
  
  
  # Model 3 excludes those w 2+ infectious symptoms at baseline or 3mo 
  cat("\nModel 3: Re-run excluding those with 2+ infectious symptoms at baseline OR 3mo------------\n")

  df_tmp_exc3 <- df_tmp %>%  
    filter(sr_3mo_curhealth_2infection ==0 & 
             sr_base_curhealth_2infection ==0)
  
  print(describeBy(df_tmp_exc3 %>% dplyr::select(contains(target_var)), group = df_tmp_exc3$sr_scr_rand_group, digits = 3))
  
  model3 <- run_model(formula_simple, df_tmp_exc3, "Model 3")
  
  cat('\n~~~~~~~~~~~~~~~~\nModel 3: ', 'dbs_3mo_', target_var, ' ~ dbs_base_', target_var, ' + sr_scr_rand_group', '\n~~~~~~~~~~~~~~~~\n')
  
  print(summary(model3))
  print(plot_residuals(model3, df_tmp, "Model 3", target_var))
  
  
  lm_model_3_d <- compute_effect_size_tobit(fit = model3,
                                            data = df_tmp_exc3 %>% filter(!is.na(!!sym(base_var)) &
                                                                            !is.na(!!sym(fu_var))),
                                            base_var = base_var,
                                            outcome_var = fu_var,
                                            left_val = left_val,
                                            ref_level = ref_level)
  
  lm_model3_results_all <- dbs_results(est = model3,
                                   target_var = target_var,
                                    smd_results =lm_model_3_d,
                                    model_name = "Excluding those with 2+ infectious symptoms at Baseline or Week 17",
                                   ref_level = ref_level)
  lm_model3_results <- lm_model3_results_all$summary_table
  lm_model3_ci <- lm_model3_results_all$ci_fixed
  model_list[["infection"]] <- model3
  
  
    # Model 4 excludes those w 2+ infectious symptoms at baseline or 3mo AND any >10 CRP 
  cat("\nModel 4: Re-run excluding CRP > 10 and those with 2+ infectious symptoms at baseline OR 3mo------------\n")

  df_tmp_exc4 <- df_tmp_exc2 %>%  
    filter(sr_3mo_curhealth_2infection == 0 & 
             sr_base_curhealth_2infection == 0)
  
  print(describeBy(df_tmp_exc4 %>% dplyr::select(contains(target_var)), group = df_tmp_exc4$sr_scr_rand_group, digits = 3))
  
  model4 <- run_model(formula_simple, df_tmp_exc4, "Model 4")
  
  cat('\n~~~~~~~~~~~~~~~~\nModel 4: ', 'dbs_3mo_', target_var, ' ~ dbs_base_', target_var, ' + sr_scr_rand_group', '\n~~~~~~~~~~~~~~~~\n')
  
  print(summary(model4))
  print(plot_residuals(model4, df_tmp, "Model 4", target_var))
  
    
  lm_model_4_d <- compute_effect_size_tobit(fit = model4,
                                            data = df_tmp_exc4 %>% filter(!is.na(!!sym(base_var)) &
                                                                            !is.na(!!sym(fu_var))),
                                            base_var = base_var,
                                            outcome_var = fu_var,
                                            left_val = left_val,
                                            ref_level = ref_level)
  
  lm_model4_results_all <- dbs_results(est = model4,
                                   target_var = target_var,
                                    smd_results =lm_model_4_d,
                                    model_name = "Excluding CRP>10 and those with 2+ infectious symptoms at Baseline or Week 17",
                                   ref_level = ref_level)
  lm_model4_results <- lm_model4_results_all$summary_table
  lm_model4_ci <- lm_model4_results_all$ci_fixed
  model_list[["crpinfection"]] <- model4
  
  result_summary <- rbind(
    lm_model1_results,
    lm_model2_results,
    lm_model3_results,
    lm_model4_results
  )
  
  ci_fixed <- list(
    Original = lm_model1_ci,
    crp = lm_model2_ci,
    infection = lm_model3_ci,
    crpinfection = lm_model4_ci
  )
  
  print(ci_fixed)
  
  return(list(result_summary = result_summary, 
              ci_fixed = ci_fixed, 
              model_list = model_list))
}
```

### Function: Simple linear regression revised for DBS data
```{r dbs-linear-regression}
lm_comparison_eh <- function(df_tmp, target_var, exc_outlier = FALSE, ref_level = 'hmp_full'){
  model_list <- list()
  df_tmp <- df_tmp %>%
    mutate(sr_scr_rand_group = relevel(factor(sr_scr_rand_group), ref = ref_level))
  
  # create baseline appending dbs_base to target
  base_var <- paste0('dbs_base_', target_var)
  # create followup (3mo) appending dbs_3mo to target 
  fu_var <- paste0('dbs_3mo_', target_var)

  if (exc_outlier == TRUE){
    mean_y <- mean(df_tmp[[fu_var]], na.rm = TRUE)
    sd_y <- sd(df_tmp[[fu_var]], na.rm = TRUE)
    threshold_upper_y <- mean_y + 3 * sd_y
    threshold_lower_y <- mean_y - 3 * sd_y
    
    mean_base <- mean(df_tmp[[base_var]], na.rm = TRUE)
    sd_base <- sd(df_tmp[[base_var]], na.rm = TRUE)
    threshold_upper_base <- mean_base + 3 * sd_base
    threshold_lower_base <- mean_base - 3 * sd_base
    
    df_tmp <- df_tmp %>%
      mutate(!!sym(fu_var) := ifelse(
               !!sym(fu_var) > threshold_upper_y | !!sym(fu_var) < threshold_lower_y,
               NA, !!sym(fu_var)),
             !!sym(base_var) := ifelse(
               !!sym(base_var) > threshold_upper_base | !!sym(base_var) < threshold_lower_base,
               NA, !!sym(base_var)))
    }



  # filter based on CRP threshold
  df_tmp_exc2 <- df_tmp |> 
    filter(dbs_base_crp_plasma_mean_AB <= 10,
           dbs_3mo_crp_plasma_mean_AB <= 10)
  threshold_string <- "values > 10 at baseline or 3mo" 

  plot_residuals <- function(model, data, model_name) {
    # Extract only the rows used in the model
    model_data <- model.frame(model)
    
    residuals_df <- data.frame(Residuals = residuals(model), Group = model_data$sr_scr_rand_group)
    
    ggplot(residuals_df, aes(x = as.factor(Group), y = Residuals, fill = as.factor(Group))) +
      geom_violin(trim = FALSE, alpha = 0.4, color = "black") +  
      geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +  
      geom_jitter(width = 0.1, size = 1, alpha = 0.6, color = "black") + 
      labs(
        title = paste("Residuals Violin Plot:", target_var, "-", model_name), 
        x = "Randomization Group", 
        y = "Residuals"
      ) +
      theme_minimal() +
      theme(
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14)
      )
  }


  cat("Running linear regression model for:", target_var, "========================================\n")
  # first run samples with infectious symptoms as covariate, ALL including outliers
  cat("Model 1: Run all samples with infectious symptoms as covariate------------\n")
  
  print(describeBy(df_tmp %>% dplyr::select(contains(target_var)), 
                   group = df_tmp$sr_scr_rand_group, digits = 3))
  
  lm_formula1a <- as.formula(paste0('dbs_3mo_', target_var, ' ~ sr_base_curhealth_2infection + sr_3mo_curhealth_2infection + dbs_base_', target_var, ' + sr_scr_rand_group'))
  lm_model1a <- lm(lm_formula1a, data = df_tmp)
  lm_model_summary1a <- summary(lm_model1a)
  
  cat('\n~~~~~~~~~~~~~~~~\nModel 1A: ', 'dbs_3mo_', target_var, ' ~ sr_base_curhealth_2infection + sr_3mo_curhealth_2infection + dbs_base_', target_var, ' + sr_scr_rand_group', '\n~~~~~~~~~~~~~~~~\n')
  
  print(lm_model_summary1a)
  print(plot_residuals(lm_model1a, df_tmp, "Model 1A"))
  
  lm_model_1a_d <- compute_effect_size_wide(estimation = lm_model1a,
                                            data = df_tmp %>% filter(!is.na(!!sym(base_var)) & !is.na(!!sym(fu_var))),
                                            base_var = base_var,
                                            outcome_var = fu_var,
                                            preMI = "Y",
                                            ref_level = ref_level)
  lm_model1a_results_all <- dbs_results(est = lm_model1a,
                                    target_var = target_var,
                                    smd_results =lm_model_1a_d,
                                    model_name = "Original",
                                    ref_level = ref_level)
  lm_model1a_results <- lm_model1a_results_all$summary_table
  lm_model1a_ci <- lm_model1a_results_all$ci_fixed
  model_list[["Original"]] <- lm_model1a
  
  # Model 2: excluding CRP>10, infectious symptoms as covariate 
  cat("\nModel 2: Re-run excluding CRP>10 with infectious symptoms as covariate------------\n")
  print(describeBy(df_tmp_exc2 %>% dplyr::select(contains(target_var)), group = df_tmp_exc2$sr_scr_rand_group, digits = 3))
  
  lm_model2a <- lm(lm_formula1a, data = df_tmp_exc2)
  lm_model_summary2a <- summary(lm_model2a)
  
  cat('\n~~~~~~~~~~~~~~~~\nModel 2A: ', 'dbs_3mo_', target_var, ' ~ sr_base_curhealth_2infection + sr_3mo_curhealth_2infection + dbs_base_', target_var, ' + sr_scr_rand_group', '\n~~~~~~~~~~~~~~~~\n')

  print(lm_model_summary2a)
  print(plot_residuals(lm_model2a, df_tmp, "Model 2A"))
  
  lm_model_2a_d <- compute_effect_size_wide(estimation = lm_model2a,
                                            data = df_tmp_exc2 %>% filter(!is.na(!!sym(base_var)) &
                                                                            !is.na(!!sym(fu_var))),
                                            base_var = base_var,
                                            outcome_var = fu_var,
                                            preMI = "Y",
                                            ref_level = ref_level)
  lm_model2a_results_all <- dbs_results(est = lm_model2a,
                                    target_var = target_var,
                                    smd_results =lm_model_2a_d,
                                    model_name = "Excluding CRP>10 with infectious symptoms as covariate",
                                    ref_level = ref_level)
  lm_model2a_results <- lm_model2a_results_all$summary_table
  lm_model2a_ci <- lm_model2a_results_all$ci_fixed
  model_list[["crp"]] <- lm_model2a 
  
  # Model 3 excludes those w 2+ infectious symptoms at baseline or 3mo 
  cat("\nModel 3: Re-run excluding those with 2+ infectious symptoms at baseline OR 3mo------------\n")

  df_tmp_exc3 <- df_tmp %>%  
    filter(sr_3mo_curhealth_2infection == 0 & 
             sr_base_curhealth_2infection == 0)
  
  print(describeBy(df_tmp_exc3 %>% dplyr::select(contains(target_var)), group = df_tmp_exc3$sr_scr_rand_group, digits = 3))
  
  lm_formula3a <- as.formula(paste0('dbs_3mo_', target_var, ' ~ dbs_base_', target_var, ' + sr_scr_rand_group'))
  lm_model3a <- lm(lm_formula3a, data = df_tmp_exc3)
  lm_model_summary3a <- summary(lm_model3a)
  
  cat('\n~~~~~~~~~~~~~~~~\nModel 3A: ', 'dbs_3mo_', target_var, ' ~ dbs_base_', target_var, ' + sr_scr_rand_group', '\n~~~~~~~~~~~~~~~~\n')
  
  print(lm_model_summary3a)
  print(plot_residuals(lm_model3a, df_tmp, "Model 3A"))
  
  lm_model_3a_d <- compute_effect_size_wide(estimation = lm_model3a,
                                            data = df_tmp_exc3 %>% filter(!is.na(!!sym(base_var)) &
                                                                            !is.na(!!sym(fu_var))),
                                            base_var = base_var,
                                            outcome_var = fu_var,
                                            preMI = "Y",
                                            ref_level = ref_level)
  
  lm_model3a_results_all <- dbs_results(est = lm_model3a,
                                    target_var = target_var,
                                    smd_results =lm_model_3a_d,
                                    model_name = "Excluding those with 2+ infectious symptoms at Baseline or Week 17",
                                    ref_level = ref_level)
  lm_model3a_results <- lm_model3a_results_all$summary_table
  lm_model3a_ci <- lm_model3a_results_all$ci_fixed
  model_list[["infection"]] <- lm_model3a
  
    # Model 4 excludes CRP>10 and those w 2+ infectious symptoms at baseline or 3mo 
  cat("\nModel 4: Re-run excluding CRP>10 and those with 2+ infectious symptoms at baseline OR 3mo------------\n")

  df_tmp_exc4 <- df_tmp_exc2 %>%  
    filter(sr_3mo_curhealth_2infection == 0 & 
             sr_base_curhealth_2infection == 0)
  
  print(describeBy(df_tmp_exc4 %>% dplyr::select(contains(target_var)), 
                   group = df_tmp_exc4$sr_scr_rand_group, digits = 3))
  
  lm_formula4a <- as.formula(paste0('dbs_3mo_', target_var, ' ~ dbs_base_', target_var, ' + sr_scr_rand_group'))
  lm_model4a <- lm(lm_formula4a, data = df_tmp_exc4)
  lm_model_summary4a <- summary(lm_model4a)
  
  cat('\n~~~~~~~~~~~~~~~~\nModel 4A: ', 'dbs_3mo_', target_var, ' ~ dbs_base_', target_var, ' + sr_scr_rand_group', '\n~~~~~~~~~~~~~~~~\n')
  
  print(lm_model_summary4a)
  print(plot_residuals(lm_model4a, df_tmp, "Model 4A"))
  
  lm_model_4a_d <- compute_effect_size_wide(estimation = lm_model4a,
                                            data = df_tmp_exc4 %>% filter(!is.na(!!sym(base_var)) &
                                                                            !is.na(!!sym(fu_var))),
                                            base_var = base_var,
                                            outcome_var = fu_var,
                                            preMI = "Y",
                                            ref_level = ref_level)
  
  lm_model4a_results_all <- dbs_results(est = lm_model4a,
                                    target_var = target_var,
                                    smd_results =lm_model_4a_d,
                                    model_name = "Excluding CRP>10 and those with 2+ infectious symptoms at Baseline or Week 17",
                                    ref_level = ref_level)
  lm_model4a_results <- lm_model4a_results_all$summary_table
  lm_model4a_ci <- lm_model4a_results_all$ci_fixed
  model_list[["crpinfection"]] <- lm_model4a
  
  result_summary <- rbind(
    lm_model1a_results,
    lm_model2a_results,
    lm_model3a_results,
    lm_model4a_results
  )
  
  ci_fixed <- list(
    Original = lm_model1a_ci,
    crp = lm_model2a_ci,
    infection = lm_model3a_ci,
    crpinfection = lm_model4a_ci
  )
  print(ci_fixed)
  
  return(list(result_summary = result_summary, 
              ci_fixed = ci_fixed, 
              model_list = model_list))
}
```

## Function: FDR correction
```{r fx_fdr_correction}
fdr_correction <- function(summary_data, group_vars = c('var_prefix', 'model_type','comparison', 'timepoint')){
  fdr_corrected_p <- summary_data %>% 
    mutate(var_prefix = sapply(outcome, map_prefix)) %>%
    group_by(across(all_of(group_vars))) %>%    
    mutate(p_value_fdr = {
      idx <- outcome != "phq8_sum"
      fdr_vals <- rep(NA_real_, n())
      fdr_vals[idx] <- p.adjust(p_value[idx], method = "fdr")
      fdr_vals[!idx] <- p_value[!idx]   # PHQ8 keeps original p
      fdr_vals
    }) %>%
    ungroup()
  return(fdr_corrected_p)
}
```


# Load data

## Load df
```{r load_df}
df <- read.csv(file = './Data/BeWell_primaryPaper_deid.csv', na.strings = c("NaN", "", "NA")) %>%
  rename(id = rand_id) %>%
  mutate(sr_scr_rand_group =  factor(sr_scr_rand_group,
                                 levels = c('hmp_full', 'hmp_active', 'uc'))) %>% 
  mutate(gender_combined = recode_values(gender_combined,
                                      'Woman' ~ 0,
                                      'Man' ~ 1,
                                      'Other' ~ 2
                                      ),
         race_nhwhite = recode_values(race_nhwhite,
                                      'No' ~ 0,
                                      'Yes' ~ 1),
         education_level = recode_values(education_level,
                                         "not college graduate" ~ 0,
                                         "college graduate" ~ 1),
         income_level = recode_values(income_level,
                                      "below median" ~ 0,
                                      "above median" ~ 1)) %>% 
  mutate(gender_combined = factor(gender_combined,
                                  levels = 0:2,
                                  labels = c('Woman', 
                                             'Man',
                                             'Other')),
         race_nhwhite = factor(race_nhwhite,
                               levels = 0:1,
                               labels = c('No',
                                          'Yes')),
         education_level = factor(education_level,
                                  levels = 0:1,
                                  labels = c('not_college_graduate',
                                             'college_graduate')),
         income_level = factor(income_level,
                               levels = 0:1,
                               labels = c('below_median',
                                          'above_median'))) %>% 
  rename_with(
    ~ str_replace(., 
                  "^vlg_(.*?)_facial_(.*)$", 
                  "vlg_\\1_vlg_\\2"),
    .cols = matches("^vlg_.*_facial_")
  ) %>% 
  mutate(sr_base_medTherapy_bin = case_when(
    rowSums(across(contains('_base_medTherapy')), na.rm = FALSE) > 0 ~ 1, 
    rowSums(across(contains('_base_medTherapy')), na.rm = FALSE) == 0 ~ 0,
    TRUE ~ NA),
    sr_pw4_medTherapy_bin = case_when(
    rowSums(across(contains('_pw4_medTherapy')), na.rm = FALSE) > 0 ~ 1, 
    rowSums(across(contains('_pw4_medTherapy')), na.rm = FALSE) == 0 ~ 0,
    TRUE ~ NA),
    sr_3mo_medTherapy_bin = case_when(
    rowSums(across(contains('_3mo_medTherapy')), na.rm = FALSE) > 0 ~ 1, 
    rowSums(across(contains('_3mo_medTherapy')), na.rm = FALSE) == 0 ~ 0,
    TRUE ~ NA)
    )

```

Update df with infectious symptoms variables
```{r dbs-covariates, message = FALSE}
base_covariates <- c('sr_base_curhealth_coldflu',
                     'sr_base_curhealth_fever',
                     'sr_base_curhealth_nightsweat',
                     'sr_base_curhealth_gi',
                     'sr_base_curhealth_bloodyStool',
                     'sr_base_curhealth_frequentPee',
                     'sr_base_curhealth_rashy',
                     'sr_base_curhealth_other')

fu_covariates <- c('sr_3mo_curhealth_coldflu',
                   'sr_3mo_curhealth_fever',
                   'sr_3mo_curhealth_nightsweat',
                   'sr_3mo_curhealth_gi',
                   'sr_3mo_curhealth_bloodyStool',
                   'sr_3mo_curhealth_frequentPee',
                   'sr_3mo_curhealth_rashy',
                   'sr_3mo_curhealth_other')

df <- df |> 
  # infection counts at baseline and 3mo: 
  mutate(sr_base_curhealth_infectionCount = 
           rowSums(across(all_of(base_covariates)), na.rm = TRUE),
         sr_3mo_curhealth_infectionCount = 
           rowSums(across(all_of(fu_covariates)), na.rm = TRUE)) |> 
  
  # binary vars for infection symp: 
  mutate(sr_base_curhealth_infection = 
           if_else(sr_base_curhealth_infectionCount >= 1, 1, 
                   sr_base_curhealth_infectionCount),
         sr_3mo_curhealth_infection = 
           if_else(sr_3mo_curhealth_infectionCount >= 1, 1, 
                   sr_3mo_curhealth_infectionCount)) |> 
  
  # per Thom's advice, binary var for those w 2+ baseline infec symp:
  mutate(sr_base_curhealth_2infection = 
           case_when(sr_base_curhealth_infectionCount >= 2 ~ 1,
                     sr_base_curhealth_infectionCount < 2 ~ 0, 
                     TRUE ~ sr_base_curhealth_infectionCount),
         sr_3mo_curhealth_2infection = 
           case_when(sr_3mo_curhealth_infectionCount >= 2 ~ 1,
                     sr_3mo_curhealth_infectionCount < 2 ~ 0, 
                     TRUE ~ sr_3mo_curhealth_infectionCount))

```

## rand_ids to be excluded from DBS analysis
Exclude due to tech error
```{r dbs_exclude_ids}
dbs_exc_ids <- c('R803211',
                 'R769299',
                 'R982843')
```


## Converting df to long format
Creating general variable list

```{r create_var_list}
outcome_vars <- c('phq8_sum', 
                  'gad7_total',
                  'promisSleep_tScore',
                  'flourishing_flourishing')

dbs_vars <- c('crp_plasma_mean_AB',
              'cyto_il6_dbs_mean_AB')

mib_vars <- c('shannon_genus_estimate',
              'Enterobacteriaceae_clr',
              'Desulfovibrio_clr',
              'Faecalibacterium_clr',
              'Clostridium_IV_clr')

mst_vars <- c('mst_LDI',
             'mst_correctedRecognition')

vlg_vars <- c('linguistic_negative_sentiment',
              'linguistic_i')

beh_vars <- c(mst_vars, 
              vlg_vars)


piecewise_vars <- c(outcome_vars,
                    vlg_vars)
```

Creating baseline variable list

```{r create_base_var_list}
sr_base_vars <- character()
sr_pw1_vars <- character()
sr_pw2_vars <- character()
sr_pw3_vars <- character()
sr_post_vars <- character()
sr_3mo_vars <- character()
for (i in seq_along(outcome_vars)) {
  sr_base_vars[i] <- paste0('sr_base_', outcome_vars[i])
  sr_pw1_vars[i] <- paste0('sr_pw1_', outcome_vars[i])
  sr_pw2_vars[i] <- paste0('sr_pw2_', outcome_vars[i])
  sr_pw3_vars[i] <- paste0('sr_pw3_', outcome_vars[i])
  sr_post_vars[i] <- paste0('sr_pw4_', outcome_vars[i])
  sr_3mo_vars[i] <- paste0('sr_3mo_', outcome_vars[i])
}
sr_all <- c(sr_base_vars, sr_pw1_vars, sr_pw2_vars, sr_pw3_vars, sr_post_vars, sr_3mo_vars)

dbs_base_vars <- character()
dbs_3mo_vars <- character()
for (i in seq_along(dbs_vars)) {
  dbs_base_vars[i] <- paste0('dbs_base_', dbs_vars[i])
  dbs_3mo_vars[i] <- paste0('dbs_3mo_', dbs_vars[i])
}
dbs_all <- c(dbs_base_vars, dbs_3mo_vars)

mib_base_vars <- character()
mib_3mo_vars <- character()
for (i in seq_along(mib_vars)) {
  mib_base_vars[i] <- paste0('mib_base_', mib_vars[i])
  mib_3mo_vars[i] <- paste0('mib_3mo_', mib_vars[i])
}
mib_all <- c(mib_base_vars, mib_3mo_vars)

vlg_base_vars <- character()
vlg_pw1_vars <- character()
vlg_pw2_vars <- character()
vlg_pw3_vars <- character()
vlg_post_vars <- character()
vlg_3mo_vars <- character()
for (i in seq_along(vlg_vars)) {
  vlg_base_vars[i] <- paste0('vlg_base_', vlg_vars[i])
  vlg_pw1_vars[i] <- paste0('vlg_pw1_', vlg_vars[i])
  vlg_pw2_vars[i] <- paste0('vlg_pw2_', vlg_vars[i])
  vlg_pw3_vars[i] <- paste0('vlg_pw3_', vlg_vars[i])
  vlg_post_vars[i] <- paste0('vlg_pw4_', vlg_vars[i])
  vlg_3mo_vars[i] <- paste0('vlg_3mo_', vlg_vars[i])
}
vlg_all <- c(vlg_base_vars, vlg_pw1_vars, vlg_pw2_vars, vlg_pw3_vars, vlg_post_vars, vlg_3mo_vars)

mst_base_vars <- character()
mst_post_vars <- character()
mst_3mo_vars <- character()
for (i in seq_along(mst_vars)) {
  mst_base_vars[i] <- paste0('beh_base_', mst_vars[i])
  mst_post_vars[i] <- paste0('beh_pw4_', mst_vars[i])
  mst_3mo_vars[i] <- paste0('beh_3mo_', mst_vars[i])
}
mst_all <- c(mst_base_vars, mst_post_vars, mst_3mo_vars)

```

Generating df_long

```{r long_to_wide}

df_long_keep_hmp_vars <- as.character(df %>% 
                                     dplyr::select(((starts_with('hmp_intervention') |
                                                       starts_with('hmp_followup')) &
                                                      -ends_with('_time'))) %>% 
                                     names())

df_long_keep_vars <- c('sr_scr_rand_group',
                       'sr_scr_scr_age',
                       'gender_combined',
                       'race_nhwhite',
                       'education_level',
                       'income_level',
                       df_long_keep_hmp_vars,
                       'sr_scr_phq9_sum')


df_long <- convert_long_format(df_tmp = df, 
                               target_var_strings = piecewise_vars,
                               constant_vars = df_long_keep_vars)

```

# Tables
## Demographics
```{r demographics}
df_dem <- df %>%
  dplyr::select(sr_scr_scr_age,
                gender_combined,
                race,
                dem_base_dem_ethnicity,
                race_nhwhite,
                education_level,
                income_level,
                sr_scr_rand_group,
                sr_scr_phq9_sum,
                sr_scr_scidPast_majorDepEpisode,
                contains('majDepEpisode'),
                ends_with('_depDisorder'),
                sr_base_medTherapy_bin) %>%
  mutate(across((where(is.factor) | where(is.character) | sr_base_medTherapy_bin), ~ifelse(is.na(.), "Not Reported", as.character(.)))) %>% 
  mutate(sr_scr_rand_group = factor(sr_scr_rand_group,
                                    levels = c('hmp_full',
                                               'hmp_active',
                                               'uc'),
                                    labels = c('HMP-full',
                                               'HMP-Psychoed',
                                               'Usual Care')),
         gender_combined = factor(gender_combined,
                                  levels = c('Man',
                                             'Woman',
                                             'Other',
                                             'Not Reported')),
         race = factor(race,
                       levels = c('White',
                                 'Black',
                                 'Native',
                                 'Asian',
                                 'Hawaiian',
                                 'more_than_one',
                                 'Not Reported'),
                       labels = c('White',
                                  'Black or African American',
                                  'American Indian or Alaska Native',
                                  'Asian or Asian American',
                                  'Native Hawaiian or Other Pacific Islander',
                                  'Multi-racial',
                                 'Not Reported')),
         dem_base_dem_ethnicity = factor(dem_base_dem_ethnicity,
                                         levels = c('yes','no','info_not_provided',
                                 'Not Reported'),
                                         labels = c('Yes',
                                                    'No',
                                                    'Did not wish to provide this information',
                                 'Not Reported')),
         race_nhwhite = factor(race_nhwhite,
                               levels = c('Yes',
                                          'No',
                                          'Not Reported')),
         education_level = factor(education_level,
                                  levels = c('not_college_graduate',
                                             'college_graduate',
                                 'Not Reported'),
                                  labels = c('Not college graduate',
                                             'College Graduate',
                                 'Not Reported')),
         income_level = factor(income_level,
                               levels = c('below_median',
                                          'above_median',
                                 'Not Reported'),
                               labels = c('Below Median',
                                          'Above Median',
                                 'Not Reported')),
         sr_scr_scidPast_majorDepEpisode = factor(sr_scr_scidPast_majorDepEpisode,
                                                  levels = 0:1,
                                                  labels = c('No', 'Yes')),
         sr_scr_scidCurrent_majDepEpisode = factor(sr_scr_scidCurrent_majDepEpisode,
                                                  levels = 0:1,
                                                  labels = c('No', 'Yes')),
         scid_lifetime_majDepEpisode = factor(scid_lifetime_majDepEpisode,
                                                levels = 0:1,
                                                labels = c('No', 'Yes')),
         scid_current_depDisorder = factor(scid_current_depDisorder,
                                          levels = 0:1,
                                          labels = c('No', 'Yes')),
         scid_past_depDisorder = factor(scid_past_depDisorder,
                                        levels = 0:1,
                                        labels = c('No', 'Yes')),
         scid_lifetime_depDisorder = factor(scid_lifetime_depDisorder,
                                            levels = 0:1,
                                            labels = c('No', 'Yes')),
         sr_base_medTherapy_bin = factor(sr_base_medTherapy_bin,
                                          levels = c(1, 0, 'Not Reported'),
                                          labels = c('Yes', 'No', 'Not Reported'))) %>% 
  rename('Age' = sr_scr_scr_age,
         'Gender' = gender_combined,
         'Race' = race,
         'Hispanic' = dem_base_dem_ethnicity,
         'Non-Hispanic White' = race_nhwhite,
         'Education' = education_level,
         'Income' = income_level,
         'PHQ-9 at screening' = sr_scr_phq9_sum, 
         'Past Major Depressive Episode' = sr_scr_scidPast_majorDepEpisode,
         'Current Major Depressive Episode' = sr_scr_scidCurrent_majDepEpisode,
         'Lifetime Major Depressive Episode' = scid_lifetime_majDepEpisode,
         'Past Depressive Disorder' = scid_past_depDisorder,
         'Current Depressive Disorder' = scid_current_depDisorder,
         'Lifetime Depressive Disorder' = scid_lifetime_depDisorder,
         'Any recent mental health treatment (therapy or medication use) at Baseline' = sr_base_medTherapy_bin)

df_dem %>% 
  tbl_summary(
    by = sr_scr_rand_group,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = list(
      all_continuous() ~ 2,
      all_categorical() ~ c(0, 1)
    ),
    percent = "column",
    missing = "ifany",
  ) %>%
  add_overall() %>%
  bold_labels() %>%
  italicize_levels()
```

## Descriptive statistics of outcomes by group
```{r all_vars_desc}
title_order <- names(target_titles)
time_order <- c("Baseline", "Week 1", "Week 2", "Week 3", "Week 4", "Week 17")

desc_table_medTherapy <- df %>%
  dplyr::select(contains("medTherapy")) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  group_by(variable) %>%
  summarize(
    n = sum(!is.na(value)),
    mean = sum(value, na.rm = TRUE), # total count of ppts with that medication/therapy, naming as mean for merging purpose
    sd = sprintf("%.1f%%", mean / n * 100) # percentage of ppts with that medication/therapy, naming as sd for merging purpose
  ) %>% 
  mutate(mean = sprintf("%.0f", mean),
         median_IQR = " ") %>% 
  rename(Variable = variable)
  

desc_table_IQR <- df %>% 
  mutate(dbs_base_cyto_il6_dbs_mean_AB = ifelse(id %in% dbs_exc_ids, NA, dbs_base_cyto_il6_dbs_mean_AB),
         dbs_3mo_cyto_il6_dbs_mean_AB = ifelse(id %in% dbs_exc_ids, NA, dbs_3mo_cyto_il6_dbs_mean_AB)) %>% # exclude IL6 cases with tech errors
  dplyr::select(all_of(c(sr_all,
                         dbs_all,
                         mib_all,
                         vlg_all,
                         mst_all))) |> 
  summarise(
    across(
      everything(),
      list(
        IQR = ~ paste0(sprintf("%.2f", quantile(.x, 0.25, na.rm = TRUE)), " - ",  sprintf("%.2f", quantile(.x, 0.75, na.rm = TRUE)))
      ),
      .names = "{.col}"
    )
  ) |> 
  pivot_longer(everything(),
               names_to = "Variable",
               values_to = "IQR")
  

desc_table <- df %>% 
  mutate(dbs_base_cyto_il6_dbs_mean_AB = ifelse(id %in% dbs_exc_ids, NA, dbs_base_cyto_il6_dbs_mean_AB),
         dbs_3mo_cyto_il6_dbs_mean_AB = ifelse(id %in% dbs_exc_ids, NA, dbs_3mo_cyto_il6_dbs_mean_AB)) %>% # exclude IL6 cases with tech errors
  dplyr::select(all_of(c(sr_all,
                         dbs_all,
                         mib_all,
                         vlg_all,
                         mst_all))) %>% 
  psych::describe() %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Variable") %>%
  dplyr::select(Variable, n, mean, sd, median
                ) %>%
  mutate(mean = as.character(sprintf("%.2f", mean)),
         sd = as.character(sprintf("%.2f", sd)),
         median = as.character(sprintf("%.2f", median))) %>% 
  left_join(desc_table_IQR) |> 
  mutate(median_IQR = paste0(median, " (", IQR, ")")) %>% 
  full_join(desc_table_medTherapy) %>% 
  mutate(n = as.integer(n)) %>%
  mutate(
    Title = sapply(Variable, map_title, title_map = target_titles),
    Time = case_when(
      str_detect(Variable, '_base_') ~ 'Baseline',
      str_detect(Variable, '_pw1_') ~ 'Week 1',
      str_detect(Variable, '_pw2_') ~ 'Week 2',
      str_detect(Variable, '_pw3_') ~ 'Week 3',
      str_detect(Variable, '_pw4_') ~ 'Week 4',
      str_detect(Variable, '_3mo_') ~ 'Week 17'
    ),
    Measure = paste0(Time, ' ', Title)
  ) %>% 
  mutate(
    Title_order = factor(Title, levels = target_titles),
    Time_order = factor(Time, levels = time_order)
  ) %>%
  arrange(Title_order, Time_order) %>%
  dplyr::select(Title, Time, n, mean, sd, median_IQR
                ) %>% 
  rename('Mean/Count' = mean,
         'SD/%' = sd,
         'Median (IQR)' = median_IQR)

# By group
group_lvs <- as.character(levels(df$sr_scr_rand_group))
desc_table_grouped <- desc_table

for (group_lv in group_lvs) {
  desc_table_medTherapy_tmp <- df %>%
    filter(sr_scr_rand_group == group_lv) %>%
    dplyr::select(contains("medTherapy")) %>%
    pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
    group_by(variable) %>%
    summarize(
      n = sum(!is.na(value)),
      mean = sum(value, na.rm = TRUE), # total count of ppts with that medication/therapy, naming as mean for merging purpose
      sd = sprintf("%.1f%%", mean / n * 100) # percentage of ppts with that medication/therapy, naming as sd for merging purpose
    ) %>% 
    mutate(mean = sprintf("%.0f", mean),
         median_IQR = " ") %>% 
    rename(Variable = variable)
  
  desc_table_IQR_tmp <- df %>% 
    filter(sr_scr_rand_group == group_lv) %>%
    mutate(dbs_base_cyto_il6_dbs_mean_AB = ifelse(id %in% dbs_exc_ids, NA, dbs_base_cyto_il6_dbs_mean_AB),
           dbs_3mo_cyto_il6_dbs_mean_AB = ifelse(id %in% dbs_exc_ids, NA, dbs_3mo_cyto_il6_dbs_mean_AB)) %>% # exclude IL6 cases with tech errors
    dplyr::select(all_of(c(sr_all,
                         dbs_all,
                         mib_all,
                         vlg_all,
                         mst_all))) |> 
    summarise(
      across(
        everything(),
        list(
          IQR = ~ paste0(sprintf("%.2f", quantile(.x, 0.25, na.rm = TRUE)), " - ",  sprintf("%.2f", quantile(.x, 0.75, na.rm = TRUE)))
        ),
        .names = "{.col}"
      )
    ) |> 
    pivot_longer(everything(),
                 names_to = "Variable",
                 values_to = "IQR")
  
  desc_table_tmp <- df %>%
    mutate(dbs_base_cyto_il6_dbs_mean_AB = ifelse(id %in% dbs_exc_ids, NA, dbs_base_cyto_il6_dbs_mean_AB),
           dbs_3mo_cyto_il6_dbs_mean_AB = ifelse(id %in% dbs_exc_ids, NA, dbs_3mo_cyto_il6_dbs_mean_AB)) %>% # exclude IL6 cases with tech errors
    filter(sr_scr_rand_group == group_lv) %>%
    dplyr::select(all_of(c(sr_all,
                           dbs_all,
                           mib_all,
                           vlg_all,
                           mst_all))) %>%
    psych::describe() %>%
    as.data.frame() %>%
    tibble::rownames_to_column("Variable") %>%
    dplyr::select(Variable, n, mean, sd, median
                  ) %>%
    mutate(mean = as.character(sprintf("%.2f", mean)),
           sd = as.character(sprintf("%.2f", sd)),
         median = as.character(sprintf("%.2f", median))) %>% 
    left_join(desc_table_IQR) %>% 
    mutate(median_IQR = paste0(median, " (", IQR, ")")) %>% 
    full_join(desc_table_medTherapy_tmp) %>% 
    mutate(n = as.integer(n)) %>%
    mutate(
      Title = sapply(Variable, map_title, title_map = target_titles),
      Time = case_when(
        str_detect(Variable, '_base_') ~ 'Baseline',
        str_detect(Variable, '_pw1_') ~ 'Week 1',
        str_detect(Variable, '_pw2_') ~ 'Week 2',
        str_detect(Variable, '_pw3_') ~ 'Week 3',
        str_detect(Variable, '_pw4_') ~ 'Week 4',
        str_detect(Variable, '_3mo_') ~ 'Week 17'
      ),
      Measure = paste0(Time, ' ', Title)
    ) %>%
    mutate(
      Title_order = factor(Title, levels = target_titles),
      Time_order = factor(Time, levels = time_order)
      ) %>%
    arrange(Title_order, Time_order) %>% 
    dplyr::select(Title, Time, n, mean, sd, median_IQR
                  ) %>%
  rename('Mean' = mean,
         'SD' = sd,
         'Median (IQR)' = median_IQR) %>% 
    rename_with(~paste0(., "_", group_lv), -c(Title, Time))
  
  desc_table_grouped <- left_join(desc_table_grouped, desc_table_tmp)
}

# Build gt table
gt_table <- desc_table_grouped %>%
  gt(groupname_col = "Title") %>% 
  cols_label(Time = "Outcome/Time point") %>% 
  fmt_number(columns = matches("(^|_)n(_|$)"), decimals = 0) %>%
  fmt_number(columns = where(is.numeric) & !matches("(^|_)n(_|$)"), decimals = 2) %>%
  tab_spanner(
    label = "All Participants",
    columns = c(n, 'Mean/Count', 'SD/%', 'Median (IQR)'
                )
  )

for (group_lv in group_lvs) {
  label_list <- setNames(
    list("n", 'Mean/Count', 'SD/%', 'Median (IQR)'
         ),
    c(paste0("n_", group_lv),
      paste0("Mean_", group_lv),
      paste0("SD_", group_lv),
      paste0("Median (IQR)_", group_lv)
      )
  )
  
  if (group_lv == 'hmp_full'){
    lv_name <- 'HMP-Full'
    } else if (group_lv == "hmp_active"){
      lv_name <- 'HMP-Psychoed'
    } else if (group_lv == "uc"){
      lv_name <- 'Usual Care'
      }
  
  gt_table <- gt_table %>%
    tab_spanner(
      label = paste(lv_name),
      id = paste0("spanner_", group_lv),
      columns = c(
        paste0("n_", group_lv),
        paste0("Mean_", group_lv),
        paste0("SD_", group_lv),
        paste0("Median (IQR)_", group_lv)
      )
    ) %>%
    cols_label(.list = label_list)
}

gt_table <- gt_table %>%
    cols_align(
          align = "center",
          columns = -2
          )
gt_table
```


# Multiple imputation for long format data

```{r mi_df}
df_pre_mi <- df %>% dplyr::select(id,
                                  sr_scr_rand_group,
                                  sr_scr_scr_age, 
                                  gender_combined, 
                                  race_nhwhite, 
                                  education_level, 
                                  income_level, 
                                  all_of(c(sr_all,
                                           vlg_all
                                           )))

var_list <- df_pre_mi %>% 
  dplyr::select(-id,
                -sr_scr_rand_group,
                -sr_scr_scr_age, 
                -gender_combined, 
                -race_nhwhite, 
                -education_level, 
                -income_level) %>% 
  names()
```

Check for missingness: over 10% for all outcome variables, except at baseline.

```{r missing_check}
df_pre_mi %>% summarise(across(everything(), ~ sum(is.na(.))))

for (outcome in outcome_vars){
  df_tmp <- df_pre_mi %>% dplyr::select(ends_with(outcome))
  print(outcome)
  print(md.pattern(df_tmp))
  print(names(df_tmp))
  rm(df_tmp, outcome)
}
```

convert into long format

```{r mi}
df_mi_long_keep_vars <- c('sr_scr_rand_group',
                          'sr_scr_scr_age',
                          'gender_combined',
                          'race_nhwhite',
                          'income_level',
                          'education_level')

all_imp_vars <- c('sr_scr_scr_age',
                  'gender_combined',
                  'race_nhwhite',
                  'income_level',
                  'education_level',
                  piecewise_vars)

df_pre_mi_long <- convert_long_format(df_tmp = df_pre_mi, 
                                      target_var_strings = piecewise_vars, 
                                      constant_vars = df_mi_long_keep_vars)

df_pre_mi_long <- df_pre_mi_long %>% 
  mutate(across(all_of(all_imp_vars), ~ ifelse(is.na(.), 1, 0), .names = "missing_{.col}"))
```


# Actual Imputation

With range restriction

```{r impute}
value_range <- list("phq8_sum" = c(0, 27), 
                    "gad7_total" = c(0, 21),
                    "flourishing_flourishing" = c(0, 10),
                    "promisSleep_tScore" = c(0, 77.6),
                    "linguistic_negative_sentiment" = c(0, 100),
                    "linguistic_i" = c(0, 100)
                    )

df_mi_restricted <- imputating_missing_long(df_tmp = df_pre_mi_long, 
                                            vars = all_imp_vars, 
                                            restrict_range = "Y",
                                            range_matrix = value_range)
```



Save the completed imputed data

```{r save_imputed}
df_og_imp <- complete(df_mi_restricted, action = "long", include = TRUE)# with original data
df_imp <- complete(df_mi_restricted, action = "long", include = FALSE) # without original data
## Check number of missing values
table(is.na(df_og_imp))
table(is.na(df_imp)) # all should be FALSE
```


## Piecewise models and MNAR sensitivity analysis
Reference group = HMP-Full

Step 1: Scale the imputed values (80% and 90% / 110% and 120%) using scale_imputed_values function.
Step 2: Run the regression analysis with the original data, the imputed and scaled values. 
Step 3: Run FDR-correction

```{r piecewise_imputed_scaled}
df_og_imp_scaled <- df_og_imp
piecewise_results <- list()
piecewise_summaries <- list()

for (piecewise_var in piecewise_vars){
  #Step 1
  df_og_imp_scaled <- scale_imputed_values_long(df_og_imp_scaled, piecewise_var)
  
  #Step 2
  piecewise_results[[piecewise_var]] <- run_sensitivity_imputed_long(df_og_imp_scaled, 
                                                                     piecewise_var, 
                                                                     return_table = TRUE)
  piecewise_summary <- piecewise_results[[piecewise_var]]$summary_table
  piecewise_summaries <- dplyr::bind_rows(piecewise_summaries, piecewise_summary)
}

print(piecewise_summaries)

#Step 3
piecewise_fdr_corrected <- fdr_correction(summary_data = piecewise_summaries, 
                                          group_vars = c('var_prefix', 
                                                         'model_type',
                                                         'comparison', 
                                                         'timepoint'))
View(piecewise_fdr_corrected)
```

# Imputed data for MST
```{r mst_mi}
df_mi_keep_vars <- c('sr_scr_rand_group',
                          'sr_scr_scr_age',
                          'gender_combined',
                          'race_nhwhite',
                          'income_level',
                          'education_level')

basic_imp_vars <- c('sr_scr_scr_age',
                  'gender_combined',
                  'race_nhwhite',
                  'income_level',
                  'education_level')

df_pre_mi_mst <- df %>%
  dplyr::select(id, all_of(c(df_mi_keep_vars, mst_all))) %>%
  mutate(across(all_of(c(basic_imp_vars, mst_all)), ~ ifelse(is.na(.), 1, 0), .names = "missing_{.col}"))

df_pre_mi_mst %>% # missingness >20% at post-test and 3mo
  dplyr::select(starts_with('missing')) %>% 
  dfSummary(round.digits = 3)

df_mi_mst <-imputating_missing_wide(df_tmp = df_pre_mi_mst,
                                    vars = c(basic_imp_vars, mst_all),
                                    restrict_range = "N")

df_og_imp_mst <- complete(df_mi_mst, action = "long", include = TRUE)# with original data
df_imp_mst <- complete(df_mi_mst, action = "long", include = FALSE) # without original data
table(is.na(df_og_imp_mst))
table(is.na(df_imp_mst)) # all should be FALSE
```

## Simple linear regression and MNAR sensitivity analysis for MST
Reference group = HMP-Full

Step 1: Scale the imputed values (80% and 90% / 110% and 120%) using scale_imputed_values function.
Step 2: Run the regression analysis with the original data, the imputed and scaled values. 

MST - Use the results from the imputed model due to high level of missingness
```{r mst_imputed_scaled}
df_og_imp_mst_scaled <- df_og_imp_mst
mst_results <- list()
mst_summaries <- list()


for (mst_var in mst_vars){
  #step 1
  df_og_imp_mst_scaled <- scale_imputed_values_wide(df_og_imp_mst_scaled,
                                                    mst_var)
  
  # Step 2
  mst_results[[mst_var]] <- run_sensitivity_imputed_wide(df, 
                                                df_og_imp_mst_scaled, 
                                                mst_var,
                                                return_table = TRUE)
  
  mst_summary <- mst_results[[mst_var]]$summary_table
  mst_summaries <- dplyr::bind_rows(mst_summaries, mst_summary)
}

print(mst_summaries)

mst_fdr_corrected <- fdr_correction(summary_data = mst_summaries, 
                                    group_vars = c('var_prefix', 
                                                   'model_type',
                                                   'comparison', 
                                                   'timepoint'))
View(mst_fdr_corrected)
```

# DBS analysis
### Simple linear regression and sensitivity analysis for C-Reactive Protein
```{r crp_lm}
crp_models <- lm_comparison_eh(df, "crp_plasma_mean_AB")

crp_summary <- crp_models$result_summary %>% 
  filter(model == 'Original')


```

### Tobit regression and sensitivity analysis for IL-6
```{r il6_tobit}
tobit_models <- tobit_comparison_eh(df %>% 
                      filter(!id %in% dbs_exc_ids), # exclude likely technical errors for IL6
                    "cyto_il6_dbs_mean_AB")

tobit_summary <- tobit_models$result_summary %>% 
  filter(model == 'Original')
```

### FDR-correction for DBS results
```{r dbs_fdr_correct}

dbs_summaries <- dplyr::bind_rows(crp_summary, tobit_summary)

dbs_fdr_corrected <- fdr_correction(dbs_summaries,
                                    group_vars = c('var_prefix',
                                                   'comparison',
                                                   'timepoint'))
print(dbs_fdr_corrected)
```


# Simple linear regression and sensitivity analysis for microbiome
```{r table_dbs_sens_summary}
df <- df %>% 
  mutate(sr_base_healthHx_antibiotics_bin = case_when(sr_base_healthHx_antibiotics == 'yes' ~ 1,
                                                      sr_base_healthHx_antibiotics == 'no' ~ 0,
                                                      TRUE ~ NA))
mib_models <- list()
mib_summaries <- list()

for (mib_var in mib_vars){
  cat('\n\n===Original model===\n\n')
  mib_models[[mib_var]] <- run_lm_og_wide(outcome = mib_var, return_table = TRUE)
  mib_summary <- mib_models[[mib_var]]$results_summary
  mib_summaries <- dplyr::bind_rows(mib_summaries, mib_summary)
  
  cat('\n\n===Excluding outliers > median + 1.5xIQR===\n\n')
  run_lm_og_wide(outcome = mib_var,
                 exc_outlier = TRUE)
  
  cat('\n\n===Excluding those taking antidepressant medications at baseline===\n\n')
  run_lm_og_wide(df %>% 
                   filter(sr_base_medTherapy_medAntiDep != 1), # update file to the new filtered file
                 outcome = mib_var,
                 exc_outlier = TRUE)
  
  cat('\n\n===Excluding those taking antibiotics in the past year===\n\n')
  run_lm_og_wide(df %>% 
                   filter(sr_base_healthHx_antibiotics_bin != 1), # update file to the new filtered file
                 outcome = mib_var,
                 exc_outlier = TRUE)
}

# FDR-Correction
mib_fdr_corrected <- fdr_correction(mib_summaries,
                                    group_vars = c('var_prefix',
                                                   'comparison',
                                                   'timepoint'))
print(mib_fdr_corrected)
```

# Excluding outliers (mean +/- >3SDs)
## Piecewise model results excluding outliers (mean +/- >3SDs)
```{r exc_3sds}
exc_3sds_models <- list()
exc_3sds_summaries <- list()

# self-report and vlog
for (piecewise_var in piecewise_vars){
  exc_3sds_models[[piecewise_var]] <- run_piecewise_og_long(outcome = piecewise_var,
                                                           exc_outlier = TRUE,
                                                           return_table = TRUE)
  
  exc_3sds_summary <- exc_3sds_models[[piecewise_var]]$results_summary
  exc_3sds_summaries <- dplyr::bind_rows(exc_3sds_summaries, exc_3sds_summary)
}

# MST
for (mst_var in mst_vars){
  exc_3sds_models[[mst_var]] <- run_sensitivity_imputed_wide(df,
                                                             df_og_imp_mst_scaled,
                                                             mst_var,
                                                             exc_outlier = TRUE,
                                                             return_table = TRUE)
  
  
  exc_3sds_summary <- exc_3sds_models[[mst_var]]$summary_table
  exc_3sds_summary <- exc_3sds_summary %>% 
    filter(model_type == "Imputed")
  
  exc_3sds_summaries <- dplyr::bind_rows(exc_3sds_summaries, exc_3sds_summary)
}

# CRP
crp_exc_3sds_models <- lm_comparison_eh(df, 
                                        "crp_plasma_mean_AB",
                                        exc_outlier = TRUE)
exc_3sds_summary <- crp_exc_3sds_models$result_summary %>% 
  filter(model == 'Original')
exc_3sds_summaries <- dplyr::bind_rows(exc_3sds_summaries, exc_3sds_summary)

# IL-6
tobit_exc_3sds_models <- tobit_comparison_eh(df %>% 
                                               filter(!id %in% dbs_exc_ids), # exclude likely technical errors for IL6
                                             "cyto_il6_dbs_mean_AB",
                                             exc_outlier = TRUE)
exc_3sds_summary <- tobit_exc_3sds_models$result_summary %>% 
  filter(model == 'Original')
exc_3sds_summaries <- dplyr::bind_rows(exc_3sds_summaries, exc_3sds_summary)

# FDR-Correction----------------------------------------------------------------
exc_3sds_fdr_corrected <- fdr_correction(exc_3sds_summaries,
                                         group_vars = c('var_prefix',
                                                        'comparison',
                                                        'timepoint'))
print(n=100, exc_3sds_fdr_corrected)
```

# Changing ref group to HMP-Psychoed

### Piecewise model results (ref group = HMP-Psychoed)
```{r piecewise_hmp_active_as_ref}
ac_models <- list()
ac_summaries <- list()

# self-report and vlog
for (piecewise_var in piecewise_vars){
  ac_models[[piecewise_var]] <- run_piecewise_og_long(outcome = piecewise_var, 
                                                      ref_level = 'hmp_active',
                                                      return_table = TRUE)
  
  ac_summary <- ac_models[[piecewise_var]]$results_summary
  ac_summaries <- dplyr::bind_rows(ac_summaries, ac_summary)
}

# MST
for (mst_var in mst_vars){
  ac_models[[mst_var]] <- run_sensitivity_imputed_wide(df, 
                                                       df_og_imp_mst_scaled,
                                                       mst_var, 
                                                       ref_level = 'hmp_active',
                                                       return_table = TRUE)
  
  ac_summary <- ac_models[[mst_var]]$summary_table
  ac_summary <- ac_summary %>% 
    filter(model_type == "Imputed")
  
  ac_summaries <- dplyr::bind_rows(ac_summaries, ac_summary)
}

# CRP
crp_ac_models <- lm_comparison_eh(df, 
                                  "crp_plasma_mean_AB", 
                                  ref_level = 'hmp_active')
ac_summary <- crp_ac_models$result_summary %>% 
  filter(model == 'Original')
ac_summaries <- dplyr::bind_rows(ac_summaries, ac_summary)

# IL-6
tobit_ac_models <- tobit_comparison_eh(df %>% 
                                         filter(!id %in% dbs_exc_ids), # exclude likely technical errors for IL6
                                       "cyto_il6_dbs_mean_AB", 
                                       ref_level = 'hmp_active')
ac_summary <- tobit_ac_models$result_summary %>% 
  filter(model == 'Original')
ac_summaries <- dplyr::bind_rows(ac_summaries, ac_summary)

# Microbiome
for (mib_var in mib_vars){
  ac_models[[mib_var]] <- run_lm_og_wide(outcome = mib_var,
                                         ref_level = 'hmp_active',
                                         return_table = TRUE)
  
  ac_summary <- ac_models[[mib_var]]$results_summary
  ac_summaries <- dplyr::bind_rows(ac_summaries, ac_summary)
}

# FDR-Correction-----------------------------------------------------------
ac_fdr_corrected <- fdr_correction(ac_summaries,
                                         group_vars = c('var_prefix',
                                                        'comparison',
                                                        'timepoint'))
print(n = 100, ac_fdr_corrected)
```


# Piecewise models with PHQ9 at screening as moderator
```{r scr_phq9_moderation}
df_long <- df_long %>% 
  mutate(scr_phq9_group = case_when(sr_scr_phq9_sum < 10 ~ 0,
                                    sr_scr_phq9_sum >= 10 ~ 1,
                                    TRUE ~NA),
         scr_phq9_group = factor(scr_phq9_group,
                                 levels = 0:1,
                                 labels = c("Low PHQ-9 at screening",
                                            "High PHQ-9 at screening")))

##===PHQ-9 at screening as moderator===
run_piecewise_og_long_moderation(
        outcome = 'phq8_sum',
        og_df_long = df_long,
        moderator_var = "sr_scr_phq9_sum"
      )

##===Low PHQ-9 at screening===
run_piecewise_og_long(
        outcome = 'phq8_sum',
        og_df_long = df_long %>% 
          filter(scr_phq9_group == "Low PHQ-9 at screening")
      )


##===High PHQ-9 at screening===
run_piecewise_og_long(
        outcome = 'phq8_sum',
        og_df_long = df_long %>% 
          filter(scr_phq9_group == "High PHQ-9 at screening")
      )

```


### HMP usage
```{r hmp_usage}
# Usage by group
summary(lm(hmp_intervention_sum_days ~ group, data = df))
confint(lm(hmp_intervention_sum_days ~ group, data = df))
summary(lm(scale(hmp_intervention_sum_days) ~ group, data = df))
confint(lm(scale(hmp_intervention_sum_days) ~ group, data = df))

summary(lm(hmp_followup_sum_days ~ group, data = df))
confint(lm(hmp_followup_sum_days ~ group, data = df))
summary(lm(scale(hmp_followup_sum_days) ~ group, data = df))
confint(lm(scale(hmp_followup_sum_days) ~ group, data = df))

# Usage x time X group interaction
hmp_2way <- lmer(phq8_sum ~ time1*sr_scr_rand_group + time1*hmp_intervention_sum_days + time2*sr_scr_rand_group + time2*hmp_followup_sum_days + (time1|id), data = df_long)
summary(hmp_2way)
confint(hmp_2way)

hmp_3way <- lmer(phq8_sum ~ time1*sr_scr_rand_group*hmp_intervention_sum_days + time2*sr_scr_rand_group*hmp_followup_sum_days + (time1|id), data = df_long)
summary(hmp_3way)
confint(hmp_3way)
```

# Poisson model
```{r poisson}
phq8_levels <- df %>%
  dplyr::select(matches("phq8_sum$")) %>%
  names()

df %>%
  dplyr::select(matches("phq8_sum$")) %>%
  pivot_longer(everything(),
               names_to = "time",
               values_to = "phq8") %>%
  mutate(time = factor(time,
                       levels = phq8_levels,
                       labels = c("Baseline", "Week 1", "Week 2",
                                  "Week 3", "Week 4", "Week 17"))) %>%
  ggplot(aes(x = phq8)) +
  geom_density(fill = "grey70", color = "black", alpha = .6) +
  facet_wrap(~ time, nrow = 2) +
  labs(x = "PHQ-8", y = "Density") +
  theme_bw()
# ggsave("Data/phq8_density_by_time.png", width = 8, height = 5, dpi = 300)

df %>%
  dplyr::select("dbs_base_cyto_il6_dbs_mean_AB","dbs_3mo_cyto_il6_dbs_mean_AB") %>%
  pivot_longer(everything(),
               names_to = "time",
               values_to = "il6") %>%
  mutate(time = factor(time,
                       levels = c("dbs_base_cyto_il6_dbs_mean_AB","dbs_3mo_cyto_il6_dbs_mean_AB"),
                       labels = c("Baseline", "Week 17"))) %>%
  ggplot(aes(x = il6)) +
  geom_density(fill = "grey70", color = "black", alpha = .6) +
  facet_wrap(~ time, nrow = 2) +
  labs(x = "Interleukin-6", y = "Density") +
  theme_bw()
# ggsave("Data/il6_density_by_time.png", width = 8, height = 5, dpi = 300)

## ---- config ----------------------------------------------------------
d        <- df_long
grp      <- "sr_scr_rand_group"   # grouping/arm variable
max_score <- 24                    # PHQ-8 upper bound
wts      <- c(4, 1)                # weights on (time1:grp, time2:grp)

## ---- models ----------------------------------------------------------
rhs <- paste0("time1*", grp, " + time2*", grp, " + (time1|id)")

m.gaus <- glmmTMB(as.formula(paste("phq8_sum ~", rhs)), data = d, family = gaussian)
m.pois <- glmmTMB(as.formula(paste("phq8_sum ~", rhs)), data = d, family = poisson)

check_overdispersion(m.pois)

## ---- AIC / BIC -------------------------------------------------------
mods <- list(Gaussian = m.gaus, Poisson = m.pois)

ic <- data.frame(
  model  = names(mods),
  df     = sapply(mods, function(x) attr(logLik(x), "df")),
  logLik = sapply(mods, function(x) as.numeric(logLik(x))),
  AIC    = sapply(mods, AIC),
  BIC    = sapply(mods, BIC),
  row.names = NULL
)
print(ic[order(ic$AIC), ])

summary(m.pois)

## ---- glht equivalent: 4*(time1:grp) + 1*(time2:grp) ------------------
contrast_test <- function(model, group_term, weights = wts,
                          exponentiate = FALSE, label = NULL) {
  
  if (inherits(model, "glmmTMB")) {
    b <- glmmTMB::fixef(model)$cond; V <- vcov(model)$cond
  } else {
    b <- lme4::fixef(model);         V <- as.matrix(vcov(model))
  }
  
  tc   <- c(paste0("time1:", group_term), paste0("time2:", group_term))
  flip <- !tc %in% names(b)
  if (any(flip)) {
    alt <- c(paste0(group_term, ":time1"), paste0(group_term, ":time2"))
    tc[flip] <- alt[flip]
  }
  if (!all(tc %in% names(b)))
    stop("Not found: ", paste(setdiff(tc, names(b)), collapse = ", "),
         "\nAvailable: ", paste(names(b), collapse = ", "))
  
  K <- setNames(numeric(length(b)), names(b)); K[tc] <- weights
  K <- matrix(K, nrow = 1, dimnames = list("contrast", names(b)))
  
  est <- as.vector(K %*% b)
  se  <- as.vector(sqrt(K %*% V %*% t(K)))
  z   <- est / se
  
  out <- data.frame(
    contrast = if (is.null(label)) paste(tc, collapse = " + ") else label,
    estimate = est, se = se, z = z, p = 2 * pnorm(-abs(z)),
    ci_lower = est - qnorm(.975) * se,
    ci_upper = est + qnorm(.975) * se, row.names = NULL)
  
  if (exponentiate)
    out <- cbind(out["contrast"], ratio = exp(out$estimate),
                 r_lower = exp(out$ci_lower), r_upper = exp(out$ci_upper),
                 out[c("se", "z", "p")])
  out
}

## ---- apply across models --------------------------------------------
active <- paste0(grp, "hmp_active")
uc     <- paste0(grp, "uc")

rbind(
  contrast_test(m.gaus, active, label = "Gaussian: full vs active"),
  contrast_test(m.gaus, uc,     label = "Gaussian: full vs UC")
)

rbind(
  contrast_test(m.pois, active, exponentiate = TRUE, label = "Poisson: full vs active"),
  contrast_test(m.pois, uc,     exponentiate = TRUE, label = "Poisson: full vs UC") 
)

rbind(
  contrast_test(m.pois, active, exponentiate = FALSE, label = "Poisson: full vs active"),
  contrast_test(m.pois, uc,     exponentiate = FALSE, label = "Poisson: full vs UC") 
)


##Posterior predictive check on the marginal distribution ---
sim.g <- simulate(m.gaus, nsim = 500)
sim.p <- simulate(m.pois, nsim = 500)

obs.tab <- table(factor(dat$value, levels = 0:24)) / nrow(dat)

pp.tab <- function(simlist) {
  m <- sapply(simlist, function(y)
    table(factor(round(pmax(y, 0)), levels = 0:24)) / length(y))
  data.frame(value = 0:24,
             mean = rowMeans(m),
             lwr  = apply(m, 1, quantile, .025),
             upr  = apply(m, 1, quantile, .975))
}

pp.g <- pp.tab(sim.g); pp.p <- pp.tab(sim.p)

## observed vs predicted frequency, both models
pdf("~/Data/poisson_vs_gaussian.pdf", width = 7, height = 5) 
plot(0:24, as.numeric(obs.tab), type = "h", lwd = 3,
     xlab = "PHQ-8", ylab = "Proportion")
lines(pp.g$value, pp.g$mean, col = "red",  lwd = 2)
lines(pp.p$value, pp.p$mean, col = "blue", lwd = 2)
legend("topright", c("Observed","Gaussian","Poisson"),
       col = c("black","red","blue"), lwd = 2, bty = "n")
dev.off() 
```

